---
title: "Visualization"
---

# Library
```{r}
library(Seurat)
library(dplyr) 
library(tidyverse)
library(SCpubr)
library(ggplot2)
library(stringr)
library(progeny)
library(decoupleR)
library(cowplot)
library(patchwork)
library(ggvenn)
library(ggrepel)
library(Matrix)
library(CellChat)
library(scales)
```

# Source necessary functions
```{r}
source('R/changeClusterName.R')
source('R/addScore.R')
source('R/markerDotPlot.R')
source('R/customFeaturePlot.R')
source('R/customGeyserPlot.R')
source('R/consensusMarkers.R')
source('R/RunVarimaxPCA.R')
source('R/convertGenes_updateGenes.R')
source('R/enrichrAnnotate.R')
source('R/VlnDotPlot.R')
source('R/colorObject.R')
source('R/ec_subtypeScore.R')
```

# Load data
```{r}
# Fat microvessel data with or without ILCs merged 
fMV <- readRDS('output/fm_merge_analyzed_subset_analyzed/seurat_obj.rds')
bMV <- readRDS('output/bm_merge_analyzed_subset_analyzed/seurat_obj.rds')
neg_fMV <- readRDS('output/fm_ilc_neg_postqc_analyzed_subset_analyzed/seurat_obj.rds')
ratMV_ec <- readRDS('output/microvessel_decontx_apr2024_annotated_cleaned_endothelial_cluster.rds')
ratMV_full <- readRDS('output/microvessel_decontx_apr2024_annotated_cleaned_metadata_updated.rds')
```

# Set default parameters
```{r}
# Change Default Assay to MAGIC
DefaultAssay(fMV) <- 'MAGIC_imputed_data'
DefaultAssay(bMV) <- 'MAGIC_imputed_data'
DefaultAssay(neg_fMV) <- 'MAGIC_imputed_data'

# Change identity to vascular_subtype
fMV <- SetIdent(fMV, value = 'vascular_subtype')
bMV <- SetIdent(bMV, value = 'vascular_subtype')

# Update annotations if necessary
fMV@meta.data <- fMV@meta.data %>%
  mutate(vascular_subtype = case_when(
    vascular_subtype == 'Arterial' ~ "Capillary 3",
    vascular_subtype == 'Artery'  ~ "Arteriole",
    vascular_subtype == 'Venous'  ~ "Venule",
    vascular_subtype == 'Capillary 1'  ~ "Capillary 1",
    vascular_subtype == 'Capillary 2'  ~ "Capillary 2",
  ))

fMV@meta.data <- fMV@meta.data %>%
  mutate(dataset = case_when(
    dataset == 'fm_ilc_pos'  ~ "fMV+ILC",
    dataset == 'fm_only_pos'  ~ "fMV-Only",

  ))

# Update annotations if necessary
bMV@meta.data <- bMV@meta.data %>%
  mutate(vascular_subtype = case_when(
    vascular_subtype == 'Arterial'  ~ "Arteriole",
    vascular_subtype == 'Capillary 3'  ~ "Capillary 3",
    vascular_subtype == 'Venous'  ~ "Venule",
    vascular_subtype == 'Capillary 1'  ~ "Capillary 1",
    vascular_subtype == 'Capillary 2'  ~ "Capillary 2",
  ))

bMV@meta.data <- bMV@meta.data %>%
  mutate(dataset = case_when(
    dataset == 'bm_ilc'  ~ "bMV+ILC",
    dataset == 'bm_only'  ~ "bMV-Only",

  ))

# Update annotations if necessary
fresh_bMV@meta.data <- fresh_bMV@meta.data %>%
  mutate(vascular_subtype = case_when(
    vascular_subtype == 'Arterial'  ~ "Arteriole",
    vascular_subtype == 'Artery'  ~ "Arteriole",
    vascular_subtype == 'Venous'  ~ "Venule",
    vascular_subtype == 'Capillary 1'  ~ "Capillary 1",
    vascular_subtype == 'Capillary 2'  ~ "Capillary 2",
  ))

# Update annotations if necessary
fresh_bMV@meta.data <- fresh_bMV@meta.data %>%
  mutate(dataset = case_when(
    dataset == 'bm_fresh' ~ 'bMV-Fresh'
  ))

# Update annotations if necessary
merge@meta.data <- merge@meta.data %>%
  mutate(vascular_subtype = case_when(
    leiden_res0_25 == '0'  ~ "Venule",
    leiden_res0_25 == '1'  ~ "Arteriole",
    leiden_res0_25 == '2'  ~ "Capillary 3",
    leiden_res0_25 == '3'  ~ "Arteriole",
    leiden_res0_25 == '4'  ~ "Arteriole",
    leiden_res0_25 == '5'  ~ "Venule",
    leiden_res0_25 == '6'  ~ "Capillary 1",
    leiden_res0_25 == '7'  ~ "Venule",
    leiden_res0_25 == '8'  ~ "Venule",
    leiden_res0_25 == '9'  ~ "Capillary 2",
    leiden_res0_25 == '10'  ~ "Arteriole",
    leiden_res0_25 == '11'  ~ "Arteriole",
    leiden_res0_25 == '12'  ~ "Venule",
  ))

merge@meta.data <- merge@meta.data %>%
  mutate(dataset = case_when(
    dataset == 'bm_ilc'  ~ "bMV+ILC",
    dataset == 'bm_only'  ~ "bMV-Only",
    dataset == 'fm_ilc_pos'  ~ "fMV+ILC",
    dataset == 'fm_only_pos'  ~ "fMV-Only",
  ))

merge@meta.data <- merge@meta.data %>%
  mutate(batch = case_when(
    batch == 'output_fm_merge_analyzed_subset_analyzed'  ~ "fMV",
    batch == 'output_bm_merge_analyzed_subset_analyzed'  ~ "bMV",
  ))


all_merge@meta.data <- all_merge@meta.data %>%
  mutate(dataset = case_when(
    dataset == 'bm_ilc'  ~ "bMV+ILC",
    dataset == 'bm_only'  ~ "bMV-Only",
    dataset == 'fm_ilc_pos'  ~ "fMV+ILC",
    dataset == 'fm_only_pos'  ~ "fMV-Only",
    dataset == 'bm_fresh' ~ 'bMV-Fresh'
  ))

all_merge@meta.data <- all_merge@meta.data %>%
  mutate(batch = case_when(
    batch == 'output_fm_merge_analyzed_subset_analyzed'  ~ "fMV",
    batch == 'output_bm_merge_analyzed_subset_analyzed'  ~ "bMV",
    batch == 'output_bm_fresh_postqc_analyzed_subset_analyzed_' ~ "bMV-Fresh"
  ))

# Update annotations if necessary
all_merge@meta.data <- all_merge@meta.data %>%
  mutate(vascular_subtype = case_when(
    batch == 'fMV' & vascular_subtype == 'Arterial'    ~ "Capillary 3",
    batch == 'fMV' & vascular_subtype == 'Artery'      ~ "Arteriole",
    batch == 'fMV' & vascular_subtype == 'Venous'      ~ "Venule",
    batch == 'fMV' & vascular_subtype == 'Capillary 1' ~ "Capillary 1",
    batch == 'fMV' & vascular_subtype == 'Capillary 2' ~ "Capillary 2",
    
    batch == 'bMV' & vascular_subtype == 'Arterial'    ~ "Arteriole",
    batch == 'bMV' & vascular_subtype == 'Capillary 3' ~ "Capillary 3",
    batch == 'bMV' & vascular_subtype == 'Venous'      ~ "Venule",
    batch == 'bMV' & vascular_subtype == 'Capillary 1' ~ "Capillary 1",
    batch == 'bMV' & vascular_subtype == 'Capillary 2' ~ "Capillary 2",
    
    batch == 'bMV-Fresh' & vascular_subtype %in% c('Arterial', 'Artery') ~ "Arteriole",
    batch == 'bMV-Fresh' & vascular_subtype == 'Venous'      ~ "Venule",
    batch == 'bMV-Fresh' & vascular_subtype == 'Capillary 1' ~ "Capillary 1",
    batch == 'bMV-Fresh' & vascular_subtype == 'Capillary 2' ~ "Capillary 2",
    
    TRUE ~ vascular_subtype  # keep original if no match
  ))


# Update annotations if necessary
neg_fMV@meta.data <- neg_fMV@meta.data %>%
  mutate(cell_type = case_when(
    leiden_res0_25 == '2'  ~ "Beta",
    leiden_res0_25 == '4'  ~ "Beta",
    leiden_res0_25 == '3'  ~ "Delta",
    leiden_res0_25 == '0'  ~ "Alpha",
    leiden_res0_25 == '1'  ~ "Alpha",
    leiden_res0_25 == '5'  ~ "Alpha",
  ))


# Update annotations if necessary
ratMV_ec@meta.data <- ratMV_ec@meta.data %>%
  mutate(condition = case_when(
    condition == 'fresh' ~ "Fresh fMV",
    condition == 'graft' ~ "fMV+PP"
  ))

# Update annotations if necessary
ratMV_ec@meta.data <- ratMV_ec@meta.data %>%
  mutate(cell_type = case_when(
    cluster.minor.int == 'Artery'  ~ "Arteriole",
    cluster.minor.int == 'Arteriole'  ~ "Arteriole",
    cluster.minor.int == 'Capillary'  ~ "Capillary",
    cluster.minor.int == 'Venule'  ~ "Venule",
  ))

# Color Pallete
color_pallete <- c("Arteriole" = '#B09C85', 
                   "Capillary 3" = '#EFC000',
                   "Venule" = '#FDAF91',
                   "Capillary 1" = '#0073C2',
                   "Capillary 2" = "#EE0000",
                   "fMV+ILC" = "#EE8733",
                   "fMV-Only" = "#003C67",
                   "bMV+ILC" = "#931C12",
                   "bMV-Only" = "darkgreen",
                   'bMV-Fresh' = "#AF46B4",
                   "Alpha" = "#4682B4",
                   "Beta" = '#B44656',
                   "Delta" = '#46B462'
                   )

color_pallete2 <- c("Venule" = "#4682B4",
                    "Capillary" = "#AF46B4",
                    "Arteriole" = "#B47846",
                    "Artery" = "#4BB446",
                    "Fresh fMV" = "#B47846",
                    "fMV + PP" = '#4682B4',
                    "Fibroblasts" = "#EFC000",
                    "Lymphatic Endothelial Cells" = "#FDAF91",
                    "Lymphoid Cells" = "#4BB446",
                    "Myeloid Cells" = "#AF46B4",
                    "Perivascular Cells" = "#B44656",
                    "Vascular Endothelial Cells" = "#4682B4")
```

# Arterio-venous markers
```{r}
features <- c("Mgp", "Gja4", "Ephb4", "Efhd1", "Nrp2", "Efnb2")
fPlot1 <- customFeaturePlot(fMV, feature = features, reduction = "X_force", add_border = T, nrow = 1)  # With black outline

ggsave(fPlot1,
       filename = "figures/arterio-venous_markers_fmv.png", 
       width = 18, 
       height = 3)

features <- c("Mgp", "Gja4", "Ephb4", "Efhd1", "Nrp2", "Efnb2")
fPlot2 <- customFeaturePlot(bMV, feature = features, reduction = "X_force", add_border = T, nrow = 1)  # With black outline


ggsave(fPlot2,
       filename = "figures/arterio-venous_markers_bmv.png", 
       width = 18, 
       height = 3)

features <- c("Mgp", "Gja4", "Ephb4", "Efhd1", "Nrp2", "Efnb2")
fPlot3 <- customFeaturePlot(fresh_bMV, feature = features, reduction = "X_force", add_border = T, nrow = 1)  # With black outline


ggsave(fPlot3,
       filename = "figures/arterio-venous_markers_fresh_bmv.png", 
       width = 18, 
       height = 3)
```

# DimPlots for fMV
```{r}
# DimPlot by sample group
# DimPlot by sample group

# Split by dataset
dPlot1 <- DimPlot(fMV, 
                  group.by = 'dataset',
                  reduction = 'X_force',
                  pt.size = 0.3,
                  cols = color_pallete
) +
  labs(title = 'Merged', x = "DIM 1", y = "DIM 2") +
  theme(
    legend.text = element_text(face = "plain", size = 6, family = "Arial"),
    axis.text = element_blank(),
    axis.title = element_text(face = "plain", size = 6, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'right',
    plot.title = element_text(size = 6, face = "plain", family = "Arial")
  )

# Split by dataset
dPlot2 <- DimPlot(fMV, 
                  group.by = 'dataset',
                  split.by = 'dataset',
                  reduction = 'X_force',
                  pt.size = 0.5,
                  cols = color_pallete
) +
  labs(title = NULL, x = "DIM 1", y = "DIM 2") +
  theme(
    legend.text = element_text(face = "plain", size = 6, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "plain", size = 6, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'none',
    plot.title = element_text(size = 6, face = "plain", family = "Arial")
  )

combinedPlot <- plot_grid(dPlot2, dPlot1, ncol = 2, rel_widths = c(2.5, 2)) 

ggsave(combinedPlot,
       filename = "figures/dimplot_groupby_sample_fMV.png", 
       width = 6, 
       height = 3)


# Split by dataset
dPlot3 <- DimPlot(fMV, 
                  group.by = 'vascular_subtype',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = 'Merged', x = "DIM 1", y = "DIM 2") +
  theme(
    legend.text = element_blank(),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'right'
  )

# Split by dataset
dPlot4 <- DimPlot(fMV, 
                  group.by = 'vascular_subtype',
                  split.by = 'dataset',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = NULL, x = "DIM 1", y = "DIM 2") +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(face = "bold", size = 12, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'none'
  )

combinedPlot <- plot_grid(dPlot4, dPlot3, ncol = 2, rel_widths = c(2, 1)) 

ggsave(combinedPlot,
       filename = "figures/dimplot_groupby_vascular_subtype_fMV.png", 
       width = 11, 
       height = 4)
```

# DimPlots for bMV
```{r}
# DimPlot by sample group
# DimPlot by sample group
library(ggplot2)

# Split by dataset
dPlot1 <- DimPlot(bMV, 
                  group.by = 'dataset',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = 'Merged', x = "DIM 1", y = "DIM 2") +
  theme(
    legend.text = element_text(face = "bold", size = 12, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'right'
  )

# Split by dataset
dPlot2 <- DimPlot(bMV, 
                  group.by = 'dataset',
                  split.by = 'dataset',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = NULL, x = "DIM 1", y = "DIM 2") +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(face = "bold", size = 12, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'none'
  )

combinedPlot <- plot_grid(dPlot2, dPlot1, ncol = 2, rel_widths = c(2, 1.5)) 

ggsave(combinedPlot,
       filename = "figures/dimplot_groupby_sample_bMV.png", 
       width = 10, 
       height = 4)


# Split by dataset
dPlot3 <- DimPlot(bMV, 
                  group.by = 'vascular_subtype',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = 'Merged', x = "DIM 1", y = "DIM 2") +
  theme(
     legend.text = element_text(face = "bold", size = 12, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'right'
  )

# Split by dataset
dPlot4 <- DimPlot(bMV, 
                  group.by = 'vascular_subtype',
                  split.by = 'dataset',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = NULL, x = "DIM 1", y = "DIM 2") +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(face = "bold", size = 12, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'none'
  )

combinedPlot <- plot_grid(dPlot4, dPlot3, ncol = 2, rel_widths = c(2, 1.5)) 

ggsave(combinedPlot,
       filename = "figures/dimplot_groupby_vascular_subtype_bMV.png", 
       width = 11, 
       height = 4)
```
# DimPlot for fresh_bm and ilc_neg
```{r}
# Split by dataset
dPlot1 <- DimPlot(fresh_bMV, 
                  group.by = 'vascular_subtype',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = 'Merged', x = "DIM 1", y = "DIM 2") +
  theme(
    legend.text = element_blank(),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'none'
  )

ggsave(dPlot1,
       filename = "figures/dimplot_groupby_sample_bMV.png", 
       width = 4, 
       height = 4)

# Split by dataset
dPlot1 <- DimPlot(neg_fMV, 
                  group.by = 'cell_type',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = 'ILCs', x = "DIM 1", y = "DIM 2") +
  theme(
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'right'
  )

ggsave(dPlot1,
       filename = "figures/dimplot_groupby_sample_ILC.png", 
       width = 6, 
       height = 4)

```

# DimPlots for merge
```{r}
# Split by dataset
dPlot1 <- DimPlot(merge, 
                  group.by = 'dataset',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = 'Merged', x = "DIM 1", y = "DIM 2") +
  theme(
    legend.text = element_text(face = "bold", size = 12, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'right'
  )

# Split by dataset
dPlot2 <- DimPlot(merge, 
                  group.by = 'dataset',
                  split.by = 'dataset',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = NULL, x = "DIM 1", y = "DIM 2") +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(face = "bold", size = 12, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'none'
  )

combinedPlot <- plot_grid(dPlot2, dPlot1, ncol = 2, rel_widths = c(2.5, 1)) 

ggsave(combinedPlot,
       filename = "figures/dimplot_groupby_sample_merge.png", 
       width = 14, 
       height = 4)


# Split by dataset
dPlot3 <- DimPlot(merge, 
                  group.by = 'vascular_subtype',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = 'Merged', x = "DIM 1", y = "DIM 2") +
  theme(
    legend.text = element_text(face = "bold", size = 12, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'right'
  )

# Split by dataset
dPlot4 <- DimPlot(merge, 
                  group.by = 'vascular_subtype',
                  split.by = 'dataset',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = NULL, x = "DIM 1", y = "DIM 2") +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(face = "bold", size = 12, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'none'
  )

combinedPlot <- plot_grid(dPlot4, dPlot3, ncol = 2, rel_widths = c(2.5, 1)) 

ggsave(combinedPlot,
       filename = "figures/dimplot_groupby_vascular_subtype_merge.png", 
       width = 14, 
       height = 4)
```


# DimPlots for ILC/neg fMV
```{r}
# Split by dataset
dPlot3 <- DimPlot(neg_fMV, 
                  group.by = 'cell_type',
                  reduction = 'X_force',
                  pt.size = 1,
                  cols = color_pallete
) +
  labs(title = 'ILCs', x = "DIM 1", y = "DIM 2") +
  theme(
    legend.text = element_text(face = "bold", size = 12, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'right'
  )

ggsave(dPlot3,
       filename = "figures/dimplot_ilc_neg_fMV.png", 
       width = 4, 
       height = 4)
```

# DimPlots for fresh rat MV
```{r}
# Split by dataset
dPlot3 <- DimPlot(ratMV_ec, 
                  group.by = 'cell_type',
                  reduction = 'umap',
                  pt.size = 1,
                  cols = color_pallete2
) +
  labs(title = 'Fresh Rat MVs', x = "DIM 1", y = "DIM 2") +
  theme(
    legend.text = element_text(face = "bold", size = 12, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'right'
  )

ggsave(dPlot3,
       filename = "figures/dimplot_rat_fresh_MV_EC.png", 
       width = 4, 
       height = 4)


dPlot4 <- DimPlot(ratMV_full, 
                  group.by = 'cluster.major.int',
                  reduction = 'umap',
                  pt.size = 1,
                  cols = color_pallete2
) +
  labs(title = 'Fresh Rat MVs', x = "DIM 1", y = "DIM 2") +
  theme(
    legend.text = element_text(face = "bold", size = 12, family = 'Arial'),
    axis.text = element_blank(),
    axis.title = element_text(face = "bold", size = 12, family = "Arial"),
    axis.ticks = element_blank(),
    axis.line = element_line(),
    legend.title = element_blank(),
    legend.position = 'right'
  )

ggsave(dPlot4,
       filename = "figures/dimplot_rat_fresh_MV_full.png", 
       width = 8, 
       height = 6)

```


# fMV: Stacked bar plot of vascular subtype ratios across datasets
```{r}
# Get metadata from object and convert to data frame and order by frequency
# Get data from object and convert to dataframe
data <- as.data.frame(table(fMV@meta.data[['vascular_subtype']], fMV@meta.data[['dataset']])) %>%
  arrange(desc(Freq))

# Change colnames
colnames(data) <- c("Cluster", "Dataset", "Freq")


bPlot <- ggplot(data[], 
                aes(fill=Cluster, 
                    y=Freq, 
                    x=Dataset)) + 
  ylab("% Fraction of Total") +
  geom_bar(position="fill", 
           stat="identity", 
           color = "black", 
           show.legend = T) + 
  scale_fill_manual(values = color_pallete) +
  scale_y_continuous(labels = label_percent()) +  # Converts proportions to percentages
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, 
                                   size=6, 
                                   face="plain", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.text.y = element_text(size=6, 
                                   face="plain", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=6, 
                                   face="plain", 
                                   family = 'Arial', 
                                   color = "black"),
                legend.title = element_text(size = 6, 
                                    face = "plain", 
                                    family = 'Arial', 
                                    color = "black"),
        legend.text = element_text(size = 6, 
                                   face = "plain", 
                                   family = 'Arial', 
                                   color = "black"))

ggsave(bPlot,
       filename = "figures/barplot_object_fMV.svg", 
       width = 2.5, 
       height = 3)
```

# bMV: Stacked bar plot of vascular subtype ratios across datasets
```{r}
# Get metadata from object and convert to data frame and order by frequency
# Get data from object and convert to dataframe
data <- as.data.frame(table(bMV@meta.data[['vascular_subtype']], bMV@meta.data[['dataset']])) %>%
  arrange(desc(Freq))

# Change colnames
colnames(data) <- c("Cluster", "Dataset", "Freq")

data$Dataset <- factor(data$Dataset, levels = c("bMV-Only", "bMV+ILC"))


bPlot <- ggplot(data[], 
                aes(fill=Cluster, 
                    y=Freq, 
                    x=Dataset)) + 
  ylab("% of Total Nuclei") +
  geom_bar(position="fill", 
           stat="identity", 
           color = "black", 
           show.legend = T) + 
  scale_fill_manual(values = color_pallete) +
  scale_y_continuous(labels = label_percent()) +  # Converts proportions to percentages
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, 
                                   size=6, 
                                   face="plain", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.text.y = element_text(size=6, 
                                   face="plain", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=6, 
                                   face="plain", 
                                   family = 'Arial', 
                                   color = "black"))

ggsave(bPlot,
       filename = "figures/barplot_object_bMV.svg", 
       width = 2.5, 
       height = 3)
```


# fresh bMV: Stacked bar plot of vascular subtype ratios across datasets
```{r}
# Get metadata from object and convert to data frame and order by frequency
# Get data from object and convert to dataframe
data <- as.data.frame(table(fresh_bMV@meta.data[['vascular_subtype']], fresh_bMV@meta.data[['dataset']])) %>%
  arrange(desc(Freq))

# Change colnames
colnames(data) <- c("Cluster", "Dataset", "Freq")


bPlot <- ggplot(data[], 
                aes(fill=Cluster, 
                    y=Freq, 
                    x=Dataset)) + 
  ylab("Cell Ratios") +
  geom_bar(position="fill", 
           stat="identity", 
           color = "black", 
           show.legend = T) + 
  scale_fill_manual(values = color_pallete) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, 
                                   size=12, 
                                   face="bold", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.text.y = element_text(size=12, 
                                   face="bold", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=12, 
                                   face="bold", 
                                   family = 'Arial', 
                                   color = "black"))

ggsave(bPlot,
       filename = "figures/barplot_object_fresh_bMV.svg", 
       width = 2.5, 
       height = 4)
```


# merge: Stacked bar plot of vascular subtype ratios across datasets
```{r}
# Get metadata from object and convert to data frame and order by frequency
# Get data from object and convert to dataframe
data <- as.data.frame(table(merge@meta.data[['vascular_subtype']], merge@meta.data[['dataset']])) %>%
  arrange(desc(Freq))

# Change colnames
colnames(data) <- c("Cluster", "Dataset", "Freq")


bPlot <- ggplot(data[], 
                aes(fill=Cluster, 
                    y=Freq, 
                    x=Dataset)) + 
  ylab("Cell Ratios") +
  geom_bar(position="fill", 
           stat="identity", 
           color = "black", 
           show.legend = T) + 
  scale_fill_manual(values = color_pallete) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, 
                                   size=12, 
                                   face="bold", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.text.y = element_text(size=12, 
                                   face="bold", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=12, 
                                   face="bold", 
                                   family = 'Arial', 
                                   color = "black"))

ggsave(bPlot,
       filename = "figures/barplot_object_fmerge.svg", 
       width = 4, 
       height = 4)
```

# ILC: Stacked bar plot of vascular subtype ratios across datasets
```{r}
# Get metadata from object and convert to data frame and order by frequency
# Get data from object and convert to dataframe
data <- as.data.frame(table(neg_fMV@meta.data[['cell_type']], neg_fMV@meta.data[['dataset']])) %>%
  arrange(desc(Freq))

# Change colnames
colnames(data) <- c("Cluster", "Dataset", "Freq")


bPlot <- ggplot(data[], 
                aes(fill=Cluster, 
                    y=Freq, 
                    x=Dataset)) + 
  ylab("Cell Ratios") +
  geom_bar(position="fill", 
           stat="identity", 
           color = "black", 
           show.legend = T) + 
  scale_fill_manual(values = color_pallete) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, 
                                   size=12, 
                                   face="bold", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.text.y = element_text(size=12, 
                                   face="bold", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=12, 
                                   face="bold", 
                                   family = 'Arial', 
                                   color = "black"))

ggsave(bPlot,
       filename = "figures/barplot_object_neg_ilc.svg", 
       width = 2.5, 
       height = 4)
```

# Fresh rat MV ECs: Stacked bar plot
```{r}
# Get metadata from object and convert to data frame and order by frequency
# Get data from object and convert to dataframe
data <- as.data.frame(table(ratMV_ec@meta.data[['cell_type']], ratMV_ec@meta.data[['condition']])) %>%
  arrange(desc(Freq))

# Change colnames
colnames(data) <- c("Cluster", "Dataset", "Freq")

data$Dataset <- factor(data$Dataset, levels = c("Fresh fMV", "fMV+PP"))


bPlot <- ggplot(data[], 
                aes(fill=Cluster, 
                    y=Freq, 
                    x=Dataset)) + 
  ylab("% Fraction of Total") +
  geom_bar(position="fill", 
           stat="identity", 
           color = "black", 
           show.legend = T) + 
  scale_fill_manual(values = color_pallete2) +
  scale_y_continuous(labels = label_percent()) +  # Converts proportions to percentages
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, 
                                   size=6, 
                                   face="plain", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.text.y = element_text(size=6, 
                                   face="plain", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=6, 
                                   face="plain", 
                                   family = 'Arial', 
                                   color = "black"))

ggsave(bPlot,
       filename = "figures/barplot_object_rat_freshMV.svg", 
       width = 2.5, 
       height = 3)
```

# Fresh rat MV full: Stacked bar plot
```{r}
# Get metadata from object and convert to data frame and order by frequency
# Get data from object and convert to dataframe
data <- as.data.frame(table(ratMV_full@meta.data[['cluster.major.int']], ratMV_full@meta.data[['condition']])) %>%
  arrange(desc(Freq))

# Change colnames
colnames(data) <- c("Cluster", "Dataset", "Freq")

data$Dataset <- factor(data$Dataset, levels = c("Fresh fMV", "fMV+PP"))


bPlot <- ggplot(data[], 
                aes(fill=Cluster, 
                    y=Freq, 
                    x=Dataset)) + 
  ylab("% Fraction of Total") +
  geom_bar(position="fill", 
           stat="identity", 
           color = "black", 
           show.legend = T) + 
  scale_fill_manual(values = color_pallete2) +
  scale_y_continuous(labels = label_percent()) +  # Converts proportions to percentages
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, 
                                   size=12, 
                                   face="bold", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.text.y = element_text(size=12, 
                                   face="bold", 
                                   family = 'Arial', 
                                   color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=12, 
                                   face="bold", 
                                   family = 'Arial', 
                                   color = "black"))

ggsave(bPlot,
       filename = "figures/barplot_object_rat_freshMV.png", 
       width = 4, 
       height = 4)
```


# fMV: Module Score of Signature Genes
```{r}
object <- fMV

object <- SetIdent(object, value = 'vascular_subtype')

# Load dataframe with signatures
ec_df <- read.csv('rutils/ec_vascular_subtype_signatures_dataframe.csv')[,-1]
 
organ_ind <- grep("organ", ec_df$label)

ec_df <- ec_df[organ_ind,]
  
# Get all the labels
labels <- unique(ec_df$label)

score_names <- c()

for (i in 1:length(labels)) {

  genes <- ec_df %>%
    filter(label == labels[i]) %>%
    pull(genes)

  # Use addScore function to obtain scores using Seurat's method
  object <- addScore(object, genes, score_name = labels[i], score_type = 'ucell')

  # Get the score names
  score_names[i] <- paste0(labels[i], "_score_ucell")

}

# Get current colnames of meta.data
current_names <- colnames(object@meta.data)

# Identify the score columns
score_cols <- grep("^organ_.*_score_ucell$", current_names, value = TRUE)

# Create new names
new_names <- str_replace(score_cols, "^organ_(.*)_score_ucell$", "\\1")  # extract organ name
new_names <- str_to_title(new_names)  # capitalize organ name
new_names <- paste(new_names, "EC Signature")  # append suffix

# Apply the renaming
colnames(object@meta.data)[match(score_cols, current_names)] <- new_names


# Extract and clean metadata
df <- object@meta.data %>%
  select(score = `Pancreas EC Signature`, 
         subtype = vascular_subtype, 
         dataset)

# Order vascular_subtype by mean score (lowest to highest)
mean_order <- df %>%
  group_by(subtype) %>%
  summarise(mean_score = mean(score, na.rm = TRUE)) %>%
  arrange(mean_score) %>%  # lowest to highest
  pull(subtype)

# Set factor levels in the desired order
df$subtype <- factor(df$subtype, levels = mean_order)

# Plot
scoreVlnPlot <- ggplot(df, aes(x = subtype, y = score, fill = dataset)) +
  geom_violin(
  position = position_dodge(width = 0.8),  # not too wide, avoids next group
  trim = FALSE,
  scale = "width",
  width = 0.6  # narrower violins to fit within group
) +
  scale_fill_manual(values = color_pallete) +
  coord_flip() +
  labs(x = "Vascular Subtype", y = "Pancreas EC Signature", fill = "Dataset") +
  theme_minimal(base_family = "Arial", base_size = 12) +
  theme(
    plot.title = element_blank(),
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 1),
    axis.ticks = element_line(color = "black", size = 0.8)
  )

ggsave(scoreVlnPlot,
       filename = "figures/scoreVlnPlot_fMV.svg", 
       width = 5, 
       height = 4)

features_to_plot <- c(new_names)

object_sub <- subset(object, idents = c("Capillary 1"))

geyPlot1 <- customGeyserPlot(object_sub, features = features_to_plot, add_border = T)

ggsave(geyPlot1,
       filename = "figures/geyserplot_capillary_organ_bMV.png", 
       width = 4, 
       height = 6)
```

# bMV: Module Score of Signature Genes
```{r}
object <- bMV

object <- SetIdent(object, value = 'vascular_subtype')

DefaultAssay(object) <- 'MAGIC_imputed_data'

# Load dataframe with signatures
ec_df <- read.csv('rutils/ec_vascular_subtype_signatures_dataframe.csv')[,-1]
 
organ_ind <- grep("organ", ec_df$label)

ec_df <- ec_df[organ_ind,]
  
# Get all the labels
labels <- unique(ec_df$label)

score_names <- c()

for (i in 1:length(labels)) {

  genes <- ec_df %>%
    filter(label == labels[i]) %>%
    pull(genes)

  # Use addScore function to obtain scores using Seurat's method
  object <- addScore(object, genes, score_name = labels[i], score_type = 'seurat')

  # Get the score names
  score_names[i] <- paste0(labels[i], "_score")

}

# Get current colnames of meta.data
current_names <- colnames(object@meta.data)

# Identify the score columns
score_cols <- grep("^organ_.*_score$", current_names, value = TRUE)

# Create new names
new_names <- str_replace(score_cols, "^organ_(.*)_score$", "\\1")  # extract organ name
new_names <- str_to_title(new_names)  # capitalize organ name
new_names <- paste(new_names, "EC Signature")  # append suffix

# Apply the renaming
colnames(object@meta.data)[match(score_cols, current_names)] <- new_names


# Extract and clean metadata
df <- object@meta.data %>%
  select(score = `Pancreas EC Signature`, 
         subtype = vascular_subtype, 
         dataset)

# Order vascular_subtype by mean score (lowest to highest)
mean_order <- df %>%
  group_by(subtype) %>%
  summarise(mean_score = mean(score, na.rm = TRUE)) %>%
  arrange(mean_score) %>%  # lowest to highest
  pull(subtype)

# Set factor levels in the desired order
df$subtype <- factor(df$subtype, levels = mean_order)

# Plot
scoreVlnPlot <- ggplot(df, aes(x = subtype, y = score, fill = dataset)) +
  geom_violin(
  position = position_dodge(width = 0.8),  # not too wide, avoids next group
  trim = FALSE,
  scale = "width",
  width = 0.6  # narrower violins to fit within group
) +
  scale_fill_manual(values = color_pallete) +
  scale_color_manual(values = c("bMV-Only" = "#931C12","bMV+ILC" = "#000000")) +
  coord_flip() +
  labs(x = "Vascular Subtype", y = "Pancreas EC Signature", fill = "Dataset") +
  theme_minimal(base_family = "Arial", base_size = 12) +
  theme(
    plot.title = element_blank(),
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 1),
    axis.ticks = element_line(color = "black", size = 0.8)
  )

ggsave(scoreVlnPlot,
       filename = "figures/scoreVlnPlot_bMV.svg", 
       width = 5.5, 
       height = 4)

features_to_plot <- c(new_names)

object_sub <- subset(object, idents = c("Capillary 1"))

geyPlot1 <- customGeyserPlot(object_sub, features = features_to_plot, add_border = T)

ggsave(geyPlot1,
       filename = "figures/geyserplot_capillary_organ_bMV.png", 
       width = 5, 
       height = 6)

```

# Merge: Module Score of Signature Genes
```{r}
object <- merge

object <- SetIdent(object, value = 'vascular_subtype')

DefaultAssay(object) <- 'MAGIC_imputed_data'

# Load dataframe with signatures
ec_df <- read.csv('rutils/ec_vascular_subtype_signatures_dataframe.csv')[,-1]
 
organ_ind <- grep("organ", ec_df$label)

ec_df <- ec_df[organ_ind,]
  
# Get all the labels
labels <- unique(ec_df$label)

score_names <- c()

for (i in 1:length(labels)) {

  genes <- ec_df %>%
    filter(label == labels[i]) %>%
    pull(genes)

  # Use addScore function to obtain scores using Seurat's method
  object <- addScore(object, genes, score_name = labels[i], score_type = 'seurat')

  # Get the score names
  score_names[i] <- paste0(labels[i], "_score")

}

# Get current colnames of meta.data
current_names <- colnames(object@meta.data)

# Identify the score columns
score_cols <- grep("^organ_.*_score$", current_names, value = TRUE)

# Create new names
new_names <- str_replace(score_cols, "^organ_(.*)_score$", "\\1")  # extract organ name
new_names <- str_to_title(new_names)  # capitalize organ name
new_names <- paste(new_names, "EC Signature")  # append suffix

# Apply the renaming
colnames(object@meta.data)[match(score_cols, current_names)] <- new_names


# Extract and clean metadata
df <- object@meta.data %>%
  select(score = `Pancreas EC Signature`, 
         subtype = vascular_subtype, 
         dataset)

# Order vascular_subtype by mean score (lowest to highest)
mean_order <- df %>%
  group_by(subtype) %>%
  summarise(mean_score = mean(score, na.rm = TRUE)) %>%
  arrange(mean_score) %>%  # lowest to highest
  pull(subtype)

# Set factor levels in the desired order
df$subtype <- factor(df$subtype, levels = mean_order)

# Plot
scoreVlnPlot <- ggplot(df, aes(x = subtype, y = score, fill = dataset)) +
  geom_violin(
  position = position_dodge(width = 0.8),  # not too wide, avoids next group
  trim = FALSE,
  scale = "width",
  width = 0.6  # narrower violins to fit within group
) +
  scale_fill_manual(values = color_pallete) +
  scale_color_manual(values = color_pallete) +
  coord_flip() +
  labs(x = "Vascular Subtype", y = "Pancreas EC Signature", fill = "Dataset") +
  theme_minimal(base_family = "Arial", base_size = 12) +
  theme(
    plot.title = element_blank(),
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 1),
    axis.ticks = element_line(color = "black", size = 0.8)
  )

ggsave(scoreVlnPlot,
       filename = "figures/scoreVlnPlot_merge.png", 
       width = 7, 
       height = 5)

features_to_plot <- c(new_names)

object_sub <- subset(object, idents = c("Capillary 1"))

geyPlot1 <- customGeyserPlot(object_sub, features = features_to_plot, add_border = T)

ggsave(geyPlot1,
       filename = "figures/geyserplot_capillary_organ_merge.png", 
       width = 5, 
       height = 5)

```



# Fresh bMV: Module Score of Signature Genes
```{r}
object <- fresh_bMV

object <- SetIdent(object, value = 'vascular_subtype')

DefaultAssay(object) <- 'MAGIC_imputed_data'

# Load dataframe with signatures
ec_df <- read.csv('rutils/ec_vascular_subtype_signatures_dataframe.csv')[,-1]
 
organ_ind <- grep("organ", ec_df$label)

ec_df <- ec_df[organ_ind,]
  
# Get all the labels
labels <- unique(ec_df$label)

score_names <- c()

for (i in 1:length(labels)) {

  genes <- ec_df %>%
    filter(label == labels[i]) %>%
    pull(genes)

  # Use addScore function to obtain scores using Seurat's method
  object <- addScore(object, genes, score_name = labels[i], score_type = 'seurat')

  # Get the score names
  score_names[i] <- paste0(labels[i], "_score")

}

# Get current colnames of meta.data
current_names <- colnames(object@meta.data)

# Identify the score columns
score_cols <- grep("^organ_.*_score$", current_names, value = TRUE)

# Create new names
new_names <- str_replace(score_cols, "^organ_(.*)_score$", "\\1")  # extract organ name
new_names <- str_to_title(new_names)  # capitalize organ name
new_names <- paste(new_names, "EC Signature")  # append suffix

# Apply the renaming
colnames(object@meta.data)[match(score_cols, current_names)] <- new_names


# Extract and clean metadata
df <- object@meta.data %>%
  select(score = `Pancreas EC Signature`, 
         subtype = vascular_subtype, 
         dataset)

# Order vascular_subtype by mean score (lowest to highest)
mean_order <- df %>%
  group_by(subtype) %>%
  summarise(mean_score = mean(score, na.rm = TRUE)) %>%
  arrange(mean_score) %>%  # lowest to highest
  pull(subtype)

# Set factor levels in the desired order
df$subtype <- factor(df$subtype, levels = mean_order)

# Plot
scoreVlnPlot <- ggplot(df, aes(x = subtype, y = score, fill = dataset)) +
  geom_violin(
  position = position_dodge(width = 0.8),  # not too wide, avoids next group
  trim = FALSE,
  scale = "width",
  width = 0.6  # narrower violins to fit within group
) +
  scale_fill_manual(values = color_pallete) +
  scale_color_manual(values = c("bMV-Fresh" = "steelblue")) +
  coord_flip() +
  labs(x = "Vascular Subtype", y = "Pancreas EC Signature", fill = "Dataset") +
  theme_minimal(base_family = "Arial", base_size = 12) +
  theme(
    plot.title = element_blank(),
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 1),
    axis.ticks = element_line(color = "black", size = 0.8)
  )

ggsave(scoreVlnPlot,
       filename = "figures/scoreVlnPlot_fresh_bMV.svg", 
       width = 6, 
       height = 3.5)

features_to_plot <- c(new_names)

object_sub <- subset(object, idents = c("Capillary 1"))

geyPlot1 <- customGeyserPlot(object_sub, features = features_to_plot, add_border = T)

ggsave(geyPlot1,
       filename = "figures/geyserplot_capillary_organ_fresh_bMV.png", 
       width = 4.5, 
       height = 6)

```

# All Merge: Module Score of Signature Genes
```{r}
object <- all_merge

# Load dataframe with signatures
ec_df <- read.csv('rutils/ec_vascular_subtype_signatures_dataframe.csv')[,-1]
 
organ_ind <- grep("organ", ec_df$label)

ec_df <- ec_df[organ_ind,]
  
# Get all the labels
labels <- unique(ec_df$label)

score_names <- c()

for (i in 1:length(labels)) {

  genes <- ec_df %>%
    filter(label == labels[i]) %>%
    pull(genes)

  # Use addScore function to obtain scores using Seurat's method
  object <- addScore(object, genes, score_name = labels[i], score_type = 'seurat')

  # Get the score names
  score_names[i] <- paste0(labels[i], "_score")

}

# Get current colnames of meta.data
current_names <- colnames(object@meta.data)

# Identify the score columns
score_cols <- grep("^organ_.*_score$", current_names, value = TRUE)

# Create new names
new_names <- str_replace(score_cols, "^organ_(.*)_score$", "\\1")  # extract organ name
new_names <- str_to_title(new_names)  # capitalize organ name
new_names <- paste(new_names, "EC Signature")  # append suffix

# Apply the renaming
colnames(object@meta.data)[match(score_cols, current_names)] <- new_names


# Extract and clean metadata
df <- object@meta.data %>%
  select(score = `Pancreas EC Signature`, 
         subtype = vascular_subtype, 
         dataset)

# Order vascular_subtype by mean score (lowest to highest)
mean_order <- df %>%
  group_by(subtype) %>%
  summarise(mean_score = mean(score, na.rm = TRUE)) %>%
  arrange(mean_score) %>%  # lowest to highest
  pull(subtype)

# Set factor levels in the desired order
df$subtype <- factor(df$subtype, levels = mean_order)

# Plot
scoreVlnPlot <- ggplot(df, aes(x = subtype, y = score, fill = dataset)) +
  geom_violin(
  position = position_dodge(width = 0.8),  # not too wide, avoids next group
  trim = FALSE,
  scale = "width",
  width = 0.6  # narrower violins to fit within group
) +
  scale_fill_manual(values = color_pallete) +
  scale_color_manual(values = color_pallete) +
  coord_flip() +
  labs(x = "Vascular Subtype", y = "Pancreas EC Signature", fill = "Dataset") +
  theme_minimal(base_family = "Arial", base_size = 12) +
  theme(
    plot.title = element_blank(),
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 1),
    axis.ticks = element_line(color = "black", size = 0.8)
  )

ggsave(scoreVlnPlot,
       filename = "figures/pancreasEC_scoreVlnPlot_all_merge.svg", 
       width = 7, 
       height = 5)

# Extract and clean metadata
df <- object@meta.data %>%
  select(score = `Brain EC Signature`, 
         subtype = vascular_subtype, 
         dataset)

# Order vascular_subtype by mean score (lowest to highest)
mean_order <- df %>%
  group_by(subtype) %>%
  summarise(mean_score = mean(score, na.rm = TRUE)) %>%
  arrange(mean_score) %>%  # lowest to highest
  pull(subtype)

# Set factor levels in the desired order
df$subtype <- factor(df$subtype, levels = mean_order)

# Plot
scoreVlnPlot <- ggplot(df, aes(x = subtype, y = score, fill = dataset)) +
  geom_violin(
  position = position_dodge(width = 0.8),  # not too wide, avoids next group
  trim = FALSE,
  scale = "width",
  width = 0.6  # narrower violins to fit within group
) +
  scale_fill_manual(values = color_pallete) +
  scale_color_manual(values = color_pallete) +
  coord_flip() +
  labs(x = "Endothelial Subtype", y = "Brain EC Signature", fill = "Dataset") +
  theme_minimal(base_family = "Arial", base_size = 12) +
  theme(
    plot.title = element_blank(),
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 1),
    axis.ticks = element_line(color = "black", size = 0.8)
  )

ggsave(scoreVlnPlot,
       filename = "figures/brain_scoreVlnPlot_all_merge.svg", 
       width = 7, 
       height = 5)

```

# Fresh rat MV EC: Module Score of Signature Genes
```{r}
object <- ratMV_ec

# Load dataframe with signatures
ec_df <- read.csv('rutils/ec_vascular_subtype_signatures_dataframe.csv')[,-1]
 
organ_ind <- grep("organ", ec_df$label)

ec_df <- ec_df[organ_ind,]
  
# Get all the labels
labels <- unique(ec_df$label)

score_names <- c()

for (i in 1:length(labels)) {

  genes <- ec_df %>%
    filter(label == labels[i]) %>%
    pull(genes)

  # Use addScore function to obtain scores using Seurat's method
  object <- addScore(object, genes, score_name = labels[i], score_type = 'seurat')

  # Get the score names
  score_names[i] <- paste0(labels[i], "_score")

}

# Get current colnames of meta.data
current_names <- colnames(object@meta.data)

# Identify the score columns
score_cols <- grep("^organ_.*_score$", current_names, value = TRUE)

# Create new names
new_names <- str_replace(score_cols, "^organ_(.*)_score$", "\\1")  # extract organ name
new_names <- str_to_title(new_names)  # capitalize organ name
new_names <- paste(new_names, "EC Signature")  # append suffix

# Apply the renaming
colnames(object@meta.data)[match(score_cols, current_names)] <- new_names


# Extract and clean metadata
df <- object@meta.data %>%
  select(score = `Pancreas EC Signature`, 
         subtype = cell_type, 
         condition)

# Order vascular_subtype by mean score (lowest to highest)
mean_order <- df %>%
  group_by(subtype) %>%
  summarise(mean_score = mean(score, na.rm = TRUE)) %>%
  arrange(mean_score) %>%  # lowest to highest
  pull(subtype)

# Set factor levels in the desired order
df$subtype <- factor(df$subtype, levels = mean_order)

summary_df <- df %>%
  group_by(subtype) %>%
  summarize(
    mean = mean(score),
    Q1 = quantile(score, 0.25),
    Q3 = quantile(score, 0.75),
    .groups = 'drop'
  )


# Plot
scoreVlnPlot <- ggplot(df, aes(x = subtype, y = score, fill = subtype)) +
  geom_violin(
  position = position_dodge(width = 0.8),  # not too wide, avoids next group
  trim = FALSE,
  scale = "width",
  width = 0.6  # narrower violins to fit within group
) +
  scale_fill_manual(values = color_pallete2) +
  scale_color_manual(values = color_pallete2) +
  coord_flip() +
  labs(x = "EC Subtype", y = "Pancreas EC Signature Score", fill = "Dataset") +
  theme_minimal(base_family = "Arial", base_size = 6) +
  theme(
    plot.title = element_blank(),
    axis.text = element_text(size = 6, face = "plain", family = "Arial"),
    axis.title = element_text(size = 6, face = "plain", family = "Arial"),
    legend.text = element_text(size = 6, face = "plain", family = "Arial"),
    legend.title = element_text(size = 6, face = "plain", family = "Arial"),
    legend.position = 'none',
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 0.5),
    axis.ticks = element_line(color = "black", size = 0.5)
  )+ stat_summary(fun.y=median, geom="point", shape=18,
                 size=2, color="red")

object@active.ident <- factor(Idents(object),
                                levels = c("Arteriole", "Capillary", "Venule", "Artery"))

object@meta.data$cell_type <- factor(
  object@meta.data$cell_type,
  levels = c("Artery", "Arteriole", "Capillary", "Venule")  # your desired order
)


# Plot
scoreVlnPlot <- ggplot(df, aes(x = subtype, y = score, fill = subtype)) +
  geom_violin(
  position = position_dodge(width = 0.8),  # not too wide, avoids next group
  trim = FALSE,
  scale = "width",
  width = 0.6  # narrower violins to fit within group
) +
  
  scale_fill_manual(values = color_pallete2) +
  scale_color_manual(values = color_pallete2) +
  coord_flip() +
  labs(x = "EC Subtype", y = "Pancreas EC Signature Score", fill = "Dataset") +
  theme_minimal(base_family = "Arial", base_size = 6) +
  theme(
    plot.title = element_blank(),
    axis.text = element_text(size = 6, face = "plain", family = "Arial"),
    axis.title = element_text(size = 6, face = "plain", family = "Arial"),
    legend.text = element_text(size = 6, face = "plain", family = "Arial"),
    legend.title = element_text(size = 6, face = "plain", family = "Arial"),
    legend.position = 'none',
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 0.5),
    axis.ticks = element_line(color = "black", size = 0.5)
  )+ stat_summary(fun.y=median, geom="point", shape=18,
                 size=2, color="red")

object@active.ident <- factor(Idents(object),
                                levels = c("Arteriole", "Capillary", "Venule", "Artery"))

object@meta.data$cell_type <- factor(
  object@meta.data$cell_type,
  levels = c("Artery", "Arteriole", "Capillary", "Venule")  # your desired order
)


dPlot <- betterDotPlot(object, features = c("Pancreas EC Signature"), metadata = 'cell_type', cluster_group = F, cluster_feature = F, )[[1]]

combinedPlot <- plot_grid(scoreVlnPlot, dPlot, ncol = 2, nrow = 1, rel_widths = c(1, 1)) 

ggsave(combinedPlot,
       filename = "figures/ratMV_EC_vlnplot_dotplot_pancreas_score.svg", 
       width = 4.5, 
       height = 2)

```


# bMV: Pancreas EC signatures comparison
```{r}
object <- bMV

signature_genes_core <- convertGenes(readRDS('rutils/signature_genes.rds'), convert_to = 'mouse')
signature_genes_expanded <- convertGenes(readRDS('rutils/signature_genes_expanded.rds'), convert_to = 'mouse')
signature_genes_all <- convertGenes(readRDS('rutils/signature_genes_expanded_all.rds'), convert_to = 'mouse')

signature_genes_list <- list("core_signature" = signature_genes_core,"expanded_signature" = signature_genes_expanded, "all_signature"=signature_genes_all)
  
# Get all the labels
labels <- names(signature_genes_list)

score_names <- c()

for (i in 1:length(labels)) {

  genes <- signature_genes_list[[i]]

  # Use addScore function to obtain scores using Seurat's method
  object <- addScore(object, genes, score_name = labels[i], score_type = 'seurat')

  # Get the score names
  score_names[i] <- paste0(labels[i], "_score")

}

# Get current colnames of meta.data
current_names <- colnames(object@meta.data)

# Identify the score columns
score_cols <- grep("signature_score", current_names, value = TRUE)

for (i in 1:length(score_cols)) {q
  # Extract and clean metadata
df <- object@meta.data %>%
  select(score = score_cols[i], 
         subtype = vascular_subtype, 
         dataset)

# Order vascular_subtype by mean score (lowest to highest)
mean_order <- df %>%
  group_by(subtype) %>%
  summarise(mean_score = mean(score, na.rm = TRUE)) %>%
  arrange(mean_score) %>%  # lowest to highest
  pull(subtype)

# Set factor levels in the desired order
df$subtype <- factor(df$subtype, levels = mean_order)

# Plot
scoreVlnPlot <- ggplot(df, aes(x = subtype, y = score, fill = dataset)) +
  geom_violin(
  position = position_dodge(width = 0.8),  # not too wide, avoids next group
  trim = FALSE,
  scale = "width",
  width = 0.6  # narrower violins to fit within group
) +
  scale_fill_manual(values = color_pallete) +
  scale_color_manual(values = c("bMV-Only" = "#931C12","bMV+ILC" = "#000000")) +
  coord_flip() +
  labs(x = "Vascular Subtype", y = score_cols[i], fill = "Dataset") +
  theme_minimal(base_family = "Arial", base_size = 12) +
  theme(
    plot.title = element_blank(),
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 1),
    axis.ticks = element_line(color = "black", size = 0.8)
  )

  ggsave(scoreVlnPlot,
       filename = paste0("figures/",score_cols[i],"_scoreVlnPlot_bMV.svg"), 
       width = 5.5, 
       height = 4)
  
}

object_sub <- subset(object, idents = c("Capillary 1"))

geyPlot1 <- customGeyserPlot(object_sub, features = score_cols, add_border = T)

ggsave(geyPlot1,
       filename = "figures/geyserplot_score_comparison_bMV.png", 
       width = 3, 
       height = 5)

```

# fMV: Pancreas EC signatures comparison
```{r}
object <- fMV

signature_genes_core <- convertGenes(readRDS('rutils/signature_genes.rds'), convert_to = 'mouse')
signature_genes_expanded <- convertGenes(readRDS('rutils/signature_genes_expanded.rds'), convert_to = 'mouse')
signature_genes_all <- convertGenes(readRDS('rutils/signature_genes_expanded_all.rds'), convert_to = 'mouse')

signature_genes_list <- list("core_signature" = signature_genes_core,"expanded_signature" = signature_genes_expanded, "all_signature"=signature_genes_all)
  
# Get all the labels
labels <- names(signature_genes_list)

score_names <- c()

for (i in 1:length(labels)) {

  genes <- signature_genes_list[[i]]

  # Use addScore function to obtain scores using Seurat's method
  object <- addScore(object, genes, score_name = labels[i], score_type = 'seurat')

  # Get the score names
  score_names[i] <- paste0(labels[i], "_score")

}

# Get current colnames of meta.data
current_names <- colnames(object@meta.data)

# Identify the score columns
score_cols <- grep("signature_score", current_names, value = TRUE)

for (i in 1:length(score_cols)) {q
  # Extract and clean metadata
df <- object@meta.data %>%
  select(score = score_cols[i], 
         subtype = vascular_subtype, 
         dataset)

# Order vascular_subtype by mean score (lowest to highest)
mean_order <- df %>%
  group_by(subtype) %>%
  summarise(mean_score = mean(score, na.rm = TRUE)) %>%
  arrange(mean_score) %>%  # lowest to highest
  pull(subtype)

# Set factor levels in the desired order
df$subtype <- factor(df$subtype, levels = mean_order)

# Plot
scoreVlnPlot <- ggplot(df, aes(x = subtype, y = score, fill = dataset)) +
  geom_violin(
  position = position_dodge(width = 0.8),  # not too wide, avoids next group
  trim = FALSE,
  scale = "width",
  width = 0.6  # narrower violins to fit within group
) +
  scale_fill_manual(values = color_pallete) +
  scale_color_manual(values = c("fMV-Only" = "#931C12","fMV+ILC" = "#000000")) +
  coord_flip() +
  labs(x = "Vascular Subtype", y = score_cols[i], fill = "Dataset") +
  theme_minimal(base_family = "Arial", base_size = 12) +
  theme(
    plot.title = element_blank(),
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 1),
    axis.ticks = element_line(color = "black", size = 0.8)
  )

  ggsave(scoreVlnPlot,
       filename = paste0("figures/",score_cols[i],"_scoreVlnPlot_fMV.svg"), 
       width = 5.5, 
       height = 4)
  
}

object_sub <- subset(object, idents = c("Capillary 1"))

geyPlot1 <- customGeyserPlot(object_sub, features = score_cols, add_border = T)

ggsave(geyPlot1,
       filename = "figures/geyserplot_score_comparison_fMV.png", 
       width = 3, 
       height = 5)

```


# All Merge: Pancreas EC signatures comparison
```{r}
object <- all_merge

signature_genes_core <- convertGenes(readRDS('rutils/signature_genes.rds'), convert_to = 'mouse')
signature_genes_expanded <- convertGenes(readRDS('rutils/signature_genes_expanded.rds'), convert_to = 'mouse')
signature_genes_all <- convertGenes(readRDS('rutils/signature_genes_expanded_all.rds'), convert_to = 'mouse')

signature_genes_list <- list("core_signature" = signature_genes_core,"expanded_signature" = signature_genes_expanded, "all_signature"=signature_genes_all)
  
# Get all the labels
labels <- names(signature_genes_list)

score_names <- c()

for (i in 1:length(labels)) {

  genes <- signature_genes_list[[i]]

  # Use addScore function to obtain scores using Seurat's method
  object <- addScore(object, genes, score_name = labels[i], score_type = 'seurat')

  # Get the score names
  score_names[i] <- paste0(labels[i], "_score")

}

# Get current colnames of meta.data
current_names <- colnames(object@meta.data)

# Identify the score columns
score_cols <- grep("signature_score", current_names, value = TRUE)

for (i in 1:length(score_cols)) {q
  # Extract and clean metadata
df <- object@meta.data %>%
  select(score = score_cols[i], 
         subtype = vascular_subtype, 
         dataset)

# Order vascular_subtype by mean score (lowest to highest)
mean_order <- df %>%
  group_by(subtype) %>%
  summarise(mean_score = mean(score, na.rm = TRUE)) %>%
  arrange(mean_score) %>%  # lowest to highest
  pull(subtype)

# Set factor levels in the desired order
df$subtype <- factor(df$subtype, levels = mean_order)

# Plot
scoreVlnPlot <- ggplot(df, aes(x = subtype, y = score, fill = dataset)) +
  geom_violin(
  position = position_dodge(width = 0.8),  # not too wide, avoids next group
  trim = FALSE,
  scale = "width",
  width = 0.6  # narrower violins to fit within group
) +
  scale_fill_manual(values = color_pallete) +
  scale_color_manual(values = c("fMV-Only" = "#931C12","fMV+ILC" = "#000000")) +
  coord_flip() +
  labs(x = "Vascular Subtype", y = score_cols[i], fill = "Dataset") +
  theme_minimal(base_family = "Arial", base_size = 12) +
  theme(
    plot.title = element_blank(),
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 1),
    axis.ticks = element_line(color = "black", size = 0.8)
  )

  ggsave(scoreVlnPlot,
       filename = paste0("figures/",score_cols[i],"_scoreVlnPlot_all_merge.svg"), 
       width = 5.5, 
       height = 4)
  
}

object_sub <- subset(object, idents = c("Capillary 1"))

geyPlot1 <- customGeyserPlot(object_sub, features = score_cols, add_border = T)

ggsave(geyPlot1,
       filename = "figures/geyserplot_score_comparison_all_merge.png", 
       width = 3, 
       height = 5)
```

# CustomFeaturePlot
```{r}

fMV <- SetIdent(fMV, value = 'dataset')

fMV_ilc_only <- subset(fMV, idents = c("fMV+ILC"))

fMV_ilc_only <- SetIdent(fMV_ilc_only, value = 'vascular_subtype')

fPlot1 <- customFeaturePlot(fMV_ilc_only, feature = c('Plvap','Insr', 'Col4a1','Igfbp3', "Nkx2-3"), reduction = "X_force", add_border = T, pt_size = 1, nrow = 1, ncol = 5)  # With black outline

ggsave(fPlot1,
       filename = "figures/feature_plot_selected_signature_fMV.png", 
       width = 14, 
       height = 3)

bMV <- SetIdent(bMV, value = 'dataset')

bMV_ilc_only <- subset(bMV, idents = c("bMV+ILC"))

bMV_ilc_only <- SetIdent(bMV_ilc_only, value = 'vascular_subtype')

fPlot2 <- customFeaturePlot(bMV_ilc_only, feature = c('Plvap','Insr', 'Col4a1','Igfbp3', 'Nkx2-3'), reduction = "X_force", add_border = T, pt_size = 1, nrow = 1, ncol = 5)  # With black outline

ggsave(fPlot2,
       filename = "figures/feature_plot_selected_signature_bMV.png", 
       width = 14, 
       height = 3)


fPlot3 <- customFeaturePlot(fresh_bMV, feature = c('Plvap','Insr', 'Col4a1','Igfbp3'), reduction = "X_force", add_border = T)  # With black outline

ggsave(fPlot3,
       filename = "figures/feature_plot_selected_signature_fresh_bMV.png", 
       width = 7, 
       height = 5)

fPlot3 <- customFeaturePlot(bMV_ilc_only, feature = c('Cldn5','Slc2a1', 'Ocln','Tjp1'), reduction = "X_force", add_border = T, pt_size = 1, nrow = 1, ncol = 5)  # With black outline

ggsave(fPlot3,
       filename = "figures/feature_plot_selected_brain_genes_bMV.png", 
       width = 14, 
       height = 3)


fPlot4 <- customFeaturePlot(fresh_bMV, feature = c('Plvap','Insr', 'Col4a1','Igfbp3','Cldn5','Slc2a1', 'Ocln','Tjp1'), nrow = 2, ncol = 4, reduction = "X_force", add_border = T)  # With black outline

ggsave(fPlot4,
       filename = "figures/feature_plot_selected_brain_genes_fresh_bMV.png", 
       width = 12, 
       height = 5)

fPlot5 <- customFeaturePlot(fresh_bMV, feature = c('Plvap','Insr', 'Col4a1','Igfbp3'), reduction = "X_force", add_border = T)  # With black outline

ggsave(fPlot5,
       filename = "figures/feature_plot_selected_signature_genes_fresh_bMV.png", 
       width = 7, 
       height = 5)


fPlot6 <- customFeaturePlot(neg_fMV, feature = c('INS','GCG', 'SST'), reduction = "X_force", add_border = T, nrow = 1, ncol = 3)  # With black outline

ggsave(fPlot6,
       filename = "figures/feature_plot_selected_signature_genes_neg_fMV.png", 
       width = 9, 
       height = 3)


fPlot7 <- customFeaturePlot(fMV, feature = c('Madcam1','Cd86'), reduction = "X_force", add_border = T, nrow = 1, ncol = 3)  # With black outline

ggsave(fPlot6,
       filename = "figures/feature_plot_selected_signature_genes_neg_fMV.png", 
       width = 9, 
       height = 3)
```

# Differential Expression Analysis
```{r}

signature_genes_all <- convertGenes(readRDS('rutils/signature_genes.rds'), convert_to = 'mouse')


markers_capillary_fmv <- FindMarkers(fMV, 
                                ident.1 = c("Capillary 1")) %>%
  mutate(feature = rownames(.),
         pct.diff = pct.1 - pct.2,
         norm.pct.diff = pct.1/pct.1 - pct.2/pct.1,
         neglogpval = -log(p_val_adj),
         neglogpval = ifelse(is.infinite(neglogpval), 
                             max(-log(.$p_val_adj)[!is.infinite(-log(.$p_val_adj))]), 
                             neglogpval),
         log2FC.neglogpval = avg_log2FC*neglogpval)

markers_capillary_fmv_filt_pos <- markers_capillary_fmv %>%
  filter(p_val_adj < 0.05 & 
           pct.1 > 0.2 & 
           avg_log2FC > 1)

markers_capillary_fmv_sig_only <- markers_capillary_fmv %>%
  filter(feature %in% signature_genes_all)


  # Rename for export
  markers_capillary_fmv <- markers_capillary_fmv %>%
    rename(
      `P value` = p_val,
      `Average log2FoldChange` = avg_log2FC,
      `% Expression in capillary 1` = pct.1,
      `% Expression in all other EC subpopulations` = pct.2,
      `Bonferroni-adjusted P value` = p_val_adj,
      `% Expression Difference` = pct.diff,
      `Normalized % Expression Difference` = norm.pct.diff,
      `-Log(adjusted p value)` = neglogpval,
      `Average log2FoldChange X -Log(adjusted p value)` = log2FC.neglogpval
    )
  
    # Rename for export
  markers_capillary_fmv_filt_pos <- markers_capillary_fmv_filt_pos %>%
    rename(
      `P value` = p_val,
      `Average log2FoldChange` = avg_log2FC,
      `% Expression in capillary 1` = pct.1,
      `% Expression in all other EC subpopulations` = pct.2,
      `Bonferroni-adjusted P value` = p_val_adj,
      `% Expression Difference` = pct.diff,
      `Normalized % Expression Difference` = norm.pct.diff,
      `-Log(adjusted p value)` = neglogpval,
      `Average log2FoldChange X -Log(adjusted p value)` = log2FC.neglogpval
    )
  
    markers_capillary_fmv_sig_only <- markers_capillary_fmv_sig_only %>%
    rename(
      `P value` = p_val,
      `Average log2FoldChange` = avg_log2FC,
      `% Expression in capillary 1` = pct.1,
      `% Expression in all other EC subpopulations` = pct.2,
      `Bonferroni-adjusted P value` = p_val_adj,
      `% Expression Difference` = pct.diff,
      `Normalized % Expression Difference` = norm.pct.diff,
      `-Log(adjusted p value)` = neglogpval,
      `Average log2FoldChange X -Log(adjusted p value)` = log2FC.neglogpval
    )

  # Write to Excel
writexl::write_xlsx(
  list(
    "All Markers" = markers_capillary_fmv,
    "Filtered Positive Markers" = markers_capillary_fmv_filt_pos,
    "Signature Genes Only" = markers_capillary_fmv_sig_only
  ),
  path = "output/Supplementary Table 2.xlsx"
)

signature_genes_all <- convertGenes(readRDS('rutils/signature_genes.rds'), convert_to = 'mouse')


markers_capillary_bmv <- FindMarkers(bMV, 
                                ident.1 = c("Capillary 1")) %>%
  mutate(feature = rownames(.),
         pct.diff = pct.1 - pct.2,
         norm.pct.diff = pct.1/pct.1 - pct.2/pct.1,
         neglogpval = -log(p_val_adj),
         neglogpval = ifelse(is.infinite(neglogpval), 
                             max(-log(.$p_val_adj)[!is.infinite(-log(.$p_val_adj))]), 
                             neglogpval),
         log2FC.neglogpval = avg_log2FC*neglogpval)

markers_capillary_bmv_filt_pos <- markers_capillary_bmv %>%
  filter(p_val_adj < 0.05 & 
           pct.1 > 0.2 & 
           avg_log2FC > 1)

markers_capillary_bmv_sig_only <- markers_capillary_bmv %>%
  filter(feature %in% signature_genes_all)


  # Rename for export
  markers_capillary_bmv <- markers_capillary_bmv %>%
    rename(
      `P value` = p_val,
      `Average log2FoldChange` = avg_log2FC,
      `% Expression in capillary 1` = pct.1,
      `% Expression in all other EC subpopulations` = pct.2,
      `Bonferroni-adjusted P value` = p_val_adj,
      `% Expression Difference` = pct.diff,
      `Normalized % Expression Difference` = norm.pct.diff,
      `-Log(adjusted p value)` = neglogpval,
      `Average log2FoldChange X -Log(adjusted p value)` = log2FC.neglogpval
    )
  
    # Rename for export
  markers_capillary_bmv_filt_pos <- markers_capillary_bmv_filt_pos %>%
    rename(
      `P value` = p_val,
      `Average log2FoldChange` = avg_log2FC,
      `% Expression in capillary 1` = pct.1,
      `% Expression in all other EC subpopulations` = pct.2,
      `Bonferroni-adjusted P value` = p_val_adj,
      `% Expression Difference` = pct.diff,
      `Normalized % Expression Difference` = norm.pct.diff,
      `-Log(adjusted p value)` = neglogpval,
      `Average log2FoldChange X -Log(adjusted p value)` = log2FC.neglogpval
    )
  
    markers_capillary_bmv_sig_only <- markers_capillary_bmv_sig_only %>%
    rename(
      `P value` = p_val,
      `Average log2FoldChange` = avg_log2FC,
      `% Expression in capillary 1` = pct.1,
      `% Expression in all other EC subpopulations` = pct.2,
      `Bonferroni-adjusted P value` = p_val_adj,
      `% Expression Difference` = pct.diff,
      `Normalized % Expression Difference` = norm.pct.diff,
      `-Log(adjusted p value)` = neglogpval,
      `Average log2FoldChange X -Log(adjusted p value)` = log2FC.neglogpval
    )

  # Write to Excel
writexl::write_xlsx(
  list(
    "All Markers" = markers_capillary_bmv,
    "Filtered Positive Markers" = markers_capillary_bmv_filt_pos,
    "Signature Genes Only" = markers_capillary_bmv_sig_only
  ),
  path = "output/Supplementary Table 3.xlsx"
)

bMV_cap1_genes <- markers_capillary_bmv_filt_pos %>%
  pull(feature)

bMV_cap1_pancreas_sig <- markers_capillary_bmv_filt_pos %>%
  filter(feature %in% signature_genes_all) %>%
  pull(feature)

fMV_cap1_genes <- markers_capillary_fmv_filt_pos %>%
  pull(feature)

fMV_cap1_pancreas_sig <- markers_capillary_fmv_filt_pos %>%
  filter(feature %in% signature_genes_all) %>%
  pull(feature)


unique_sig_bMV <- setdiff(bMV_cap1_pancreas_sig, fMV_cap1_pancreas_sig)
unique_sig_fMV <- setdiff(fMV_cap1_pancreas_sig, bMV_cap1_pancreas_sig)
intersect_sig <- intersect(fMV_cap1_pancreas_sig, bMV_cap1_pancreas_sig)


# List all vectors
vecs <- list(
  "Pancreas EC signature genes" = signature_genes_all,
  "Overlap of fMV Capillary 1 markers with Pancreas EC signature genes" = fMV_cap1_pancreas_sig,
  "Overlap of bMV Capillary 1 markers with Pancreas EC signature genes" = bMV_cap1_pancreas_sig,
  "Overlap of fMV and bMV Pancreas EC signature genes" = intersect_sig,
  "Unique bMV Pancreas EC signature genes" = unique_sig_bMV,
  "Unique fMV Pancreas EC signature genes" = unique_sig_fMV
)

# Find max length
max_len <- max(sapply(vecs, length))

# Pad with NA
vecs_padded <- lapply(vecs, function(x) {
  length(x) <- max_len
  return(x)
})

# Create the data frame
df <- as.data.frame(vecs_padded, stringsAsFactors = FALSE)

# Save as CSV
write.csv(df, "output/Supplementary Table 4.csv", row.names = FALSE, na = "")


seq <- readxl::read_excel('rutils/tGFP_peaks_annotated_3kb_promoter.xlsx')

seq <- seq %>%
  filter(annotation == 'Promoter (<=1kb)')

chipseq_genes <- seq %>%
  pull(SYMBOL)

chipseq_genes <- convertGenes(chipseq_genes, convert_to = 'mouse')

sig_chipseq <- intersect(chipseq_genes, signature_genes_all)

bMV_sig_chipseq <- intersect(sig_chipseq, bMV_cap1_pancreas_sig)

fMV_sig_chipseq <- intersect(sig_chipseq, fMV_cap1_pancreas_sig)

unique_bMV_chipseq <- setdiff(bMV_sig_chipseq, fMV_sig_chipseq)
  
unique_fMV_chipseq <- setdiff(fMV_sig_chipseq, bMV_sig_chipseq)
  
intersect_sig_chipseq <- intersect(fMV_sig_chipseq, bMV_sig_chipseq)

# List all vectors
vecs <- list(
  "Pancreas EC signature genes w/ NKX2-3 binding sites" = sig_chipseq,
  "Overlap of bMV Capillary 1 markers with Pancreas EC signature genes w/ NKX2-3 binding sites" = bMV_sig_chipseq,
  "Overlap of fMV Capillary 1 markers with Pancreas EC signature genes w/ NKX2-3 binding sites" = fMV_sig_chipseq,
  "Overlap of fMV and bMV Pancreas EC signature genes w/ NKX2-3 binding sites" = intersect_sig_chipseq,
  "Unique bMV Pancreas EC signature genes w/ NKX2-3 binding sites" = unique_bMV_chipseq,
  "Unique fMV Pancreas EC signature genes w/ NKX2-3 binding sites" = unique_fMV_chipseq
)

# Find max length
max_len <- max(sapply(vecs, length))

# Pad with NA
vecs_padded <- lapply(vecs, function(x) {
  length(x) <- max_len
  return(x)
})

# Create the data frame
df <- as.data.frame(vecs_padded, stringsAsFactors = FALSE)

# Save as CSV
write.csv(df, "output/Supplementary Table 7.csv", row.names = FALSE, na = "")


# Color Pallete
color_pallete <- c("Arteriole" = '#B09C85', 
                   "Capillary 3" = '#EFC000',
                   "Venule" = '#FDAF91',
                   "Capillary 1" = '#0073C2',
                   "Capillary 2" = "#EE0000",
                   "fMV+ILC" = "#EE8733",
                   "fMV-Only" = "#003C67",
                   "bMV+ILC" = "#931C12",
                   "bMV-Only" = "darkgreen",
                   'bMV-Fresh' = "#AF46B4",
                   "Alpha" = "#4682B4",
                   "Beta" = '#B44656',
                   "Delta" = '#46B462'
                   )


# Overlap of fMV
venn_list1 <- list("fMV+ILC Capillary 1" = fMV_cap1_genes,
                  "Pancreatic EC Signature Genes" = signature_genes_all)

venn1 <- ggvenn(
  venn_list1, 
  fill_color = c("#EE8733", "#0073C2"),
  stroke_size = 0.5, 
  set_name_size = 4,
  show_percentage = F, text_size = 4, 
  )

ggsave(venn1,
       filename = "figures/overlap1.svg", 
       width = 5, 
       height = 6)


# Overlap of bMV
venn_list2 <- list("bMV+ILC Capillary 1" = bMV_cap1_genes,
                  "Pancreatic EC Signature Genes" = signature_genes_all)

venn2 <- ggvenn(
  venn_list2, 
  fill_color = c("#931C12", "#0073C2"),
  stroke_size = 0.5, 
  set_name_size = 4,
  show_percentage = F, text_size = 4, 
  )

ggsave(venn2,
       filename = "figures/overlap2.svg", 
       width = 5, 
       height = 6)

# Overlap of bMV
venn_list3 <- list("fMV+ILC Capillary 1" = fMV_cap1_genes,
                   "bMV+ILC Capillary 1" = bMV_cap1_genes,
                   "Pancreatic EC Signature Genes" = signature_genes_all)

venn3 <- ggvenn(
  venn_list3, 
  fill_color = c("#EE8733", "#931C12", "#0073C2"),
  stroke_size = 0.5, 
  set_name_size = 4,
  show_percentage = F, text_size = 4, 
  )

ggsave(venn3,
       filename = "figures/overlap3.svg", 
       width = 5, 
       height = 6)

# Overlap of bMV
venn_list4 <- list("fMV+ILC Capillary 1" = fMV_cap1_genes,
                   "Pancreatic EC Signature Genes" = signature_genes_all,
                   "Genes w/ NKX2-3 binding sites" = chipseq_genes)

venn4 <- ggvenn(
  venn_list4, 
  fill_color = c("#EE8733", "#0073C2", "#46B462"),
  stroke_size = 0.5, 
  set_name_size = 4,
  show_percentage = F, text_size = 4, 
  )

ggsave(venn4,
       filename = "figures/overlap4.svg", 
       width = 5, 
       height = 6)


# Overlap of bMV
venn_list5 <- list("bMV+ILC Capillary 1" = bMV_cap1_genes,
                   "Pancreatic EC Signature Genes" = signature_genes_all,
                   "Genes w/ NKX2-3 binding sites" = chipseq_genes)

venn5 <- ggvenn(
  venn_list5, 
  fill_color = c("#931C12", "#0073C2", "#46B462"),
  stroke_size = 0.5, 
  set_name_size = 4,
  show_percentage = F, text_size = 4, 
  )

ggsave(venn5,
       filename = "figures/overlap5.svg", 
       width = 5, 
       height = 6)

# Overlap of bMV
venn_list6 <- list("fMV+ILC Capillary 1" = fMV_cap1_genes,
                   "bMV+ILC Capillary 1" = bMV_cap1_genes,
                   "Pancreatic EC Signature Genes w/ NKX2-3 binding site" = sig_chipseq)

venn6 <- ggvenn(
  venn_list6, 
  fill_color = c("#EE8733", "#931C12", "#EFC000"),
  stroke_size = 0.5, 
  set_name_size = 4,
  show_percentage = F, text_size = 4, 
  )

ggsave(venn6,
       filename = "figures/overlap6.svg", 
       width = 5, 
       height = 6)


```


# Comparison of Capillary 1 markers individually for fMV and bMV or chipseq genes
```{r}
# Stringency Setting
#1 - All Signature Genes, All Markers > 0.5
#2 - All Signature Genes, All Markes > 1
#3 - Expanded Signature Genes, All Markers > 0.5
#4 - Expanded Signature Genes, All Markers > 1
#5 - Core signature genes, All Markers > 0.5
#6 - Core signature genes, All Markers > 1


seq <- readxl::read_excel('rutils/tGFP_peaks_annotated_3kb_promoter.xlsx')

seq <- seq %>%
  filter(annotation == 'Promoter (<=1kb)')

chipseq_genes <- seq %>%
  pull(SYMBOL)

chipseq_genes <- convertGenes(chipseq_genes, convert_to = 'mouse')
signature_genes_all <- convertGenes(readRDS('rutils/signature_genes.rds'), convert_to = 'mouse')

markers_capillary_fmv <- FindMarkers(fMV, 
                                ident.1 = c("Capillary 1"))

markers_capillary_fmv_filt <- markers_capillary_fmv %>%
  filter(p_val_adj < 0.05 & 
           pct.1 > 0.2 & 
           avg_log2FC > 1) %>%
  mutate(feature = rownames(.),
         pct.diff = pct.1 - pct.2,
         norm.pct.diff = pct.1/pct.1 - pct.2/pct.1,
         neglogpval = -log(p_val_adj),
         neglogpval = ifelse(is.infinite(neglogpval), 
                             max(-log(.$p_val_adj)[!is.infinite(-log(.$p_val_adj))]), 
                             neglogpval),
         log2FC.neglogpval = avg_log2FC*neglogpval)

  # Rename for export
  markers_capillary_fmv_filt_xl <- markers_capillary_fmv_filt %>%
    rename(
      `P value` = p_val,
      `Average log2FoldChange` = avg_log2FC,
      `% Expression in fMV+ILC` = pct.1,
      `% Expression in fMV-Only` = pct.2,
      `Bonferroni-adjusted P value` = p_val_adj,
      `% Expression Difference` = pct.diff,
      `Normalized % Expression Difference` = norm.pct.diff,
      `-Log(adjusted p value)` = neglogpval,
      `Average log2FoldChange X -Log(adjusted p value)` = log2FC.neglogpval
    )

  # Write to Excel
writexl::write_xlsx(markers_capillary_fmv_filt_xl, "output/Supplementary Table 2.xlsx")
  


markers_capillary_bmv <- FindMarkers(bMV, 
                                ident.1 = c("Capillary 1"))

markers_capillary_bmv_filt <- markers_capillary_bmv %>%
  filter(p_val_adj < 0.05 & 
           pct.1 > 0.2 & 
           avg_log2FC > 1) %>%
  mutate(feature = rownames(.),
         pct.diff = pct.1 - pct.2,
         norm.pct.diff = pct.1/pct.1 - pct.2/pct.1,
         neglogpval = -log(p_val_adj),
         neglogpval = ifelse(is.infinite(neglogpval), 
                             max(-log(.$p_val_adj)[!is.infinite(-log(.$p_val_adj))]), 
                             neglogpval),
         log2FC.neglogpval = avg_log2FC*neglogpval)

  # Rename for export
  markers_capillary_bmv_filt_xl <- markers_capillary_bmv_filt %>%
    rename(
      `P value` = p_val,
      `Average log2FoldChange` = avg_log2FC,
      `% Expression in fMV+ILC` = pct.1,
      `% Expression in fMV-Only` = pct.2,
      `Bonferroni-adjusted P value` = p_val_adj,
      `% Expression Difference` = pct.diff,
      `Normalized % Expression Difference` = norm.pct.diff,
      `-Log(adjusted p value)` = neglogpval,
      `Average log2FoldChange X -Log(adjusted p value)` = log2FC.neglogpval
    )

  # Write to Excel
writexl::write_xlsx(markers_capillary_bmv_filt_xl, "output/Supplementary Table 3.xlsx")


markers_capillary_fmv_filt_genes <- markers_capillary_fmv_filt$feature
markers_capillary_bmv_filt_genes <- markers_capillary_bmv_filt$feature

pancreas_sig_nkx_bind <- intersect(chipseq_genes, signature_genes_all)

fMV_cap1_pancreas_sig_nkx_bind <- intersect(markers_capillary_fmv_filt_genes, pancreas_sig_nkx_bind)
bMV_cap1_pancreas_sig_nkx_bind <- intersect(markers_capillary_bmv_filt_genes, pancreas_sig_nkx_bind)

fMV_bMV_cap1 <- intersect(markers_capillary_fmv_filt_genes, markers_capillary_bmv_filt_genes)

write.csv(fMV_cap1_pancreas_sig_nkx_bind, file = 'output/overlap_pancreas_sig_genes_nkx2-3_bind_fMV_cap1_markers_method1.csv')

write.csv(bMV_cap1_pancreas_sig_nkx_bind, file = 'output/overlap_pancreas_sig_genes_nkx2-3_bind_bMV_cap1_markers_method1.csv')

write.csv(fMV_bMV_cap1, file = 'output/overlap_fMV_cap1_bMV_cap1_markers_method1.csv')

# Overlap of all Endothelial Genes
venn_list <- list("fMV+ILC Capillary 1 Markers" = markers_capillary_fmv_filt_genes,
                  "bMV+ILC Capillary 1 Markers" = markers_capillary_bmv_filt_genes,
                  "Pancreas Sig Genes w/ NKX2-3 sites" = pancreas_sig_nkx_bind)

venn <- ggvenn(
  venn_list, 
  fill_color = c("#EE8733", "#931C12", "#008B45", "steelblue"),
  stroke_size = 0.5, 
  set_name_size = 4,
  show_percentage = F, text_size = 4, 
  )

ggsave(venn,
       filename = "figures/overlap_chipseq_signature_fMV_bMV_DEGs_ind.svg", 
       width = 5, 
       height = 6)


gsoa1 <- enrichAnnotateR(fMV_cap1_pancreas_sig_nkx_bind, database = 'function') %>%
    mutate(neglog_p = -log(Adjusted.P.value)) %>%
    filter(intersection_size > 2) %>%
  filter(term_size < 1000) %>%
  arrange(Combined.Score) %>%
  #slice_head(n = 20) %>% 
  mutate(Term = factor(Term, levels = Term))

write.csv(gsoa1, "output/Supplementary Table 6.csv", row.names = FALSE)


# Horizontal bar plot
gsoa1_plot <- ggplot(gsoa1, aes(x = Combined.Score,, y = Term)) +
  geom_point(aes(size = gene_ratio, color = neglog_p, fill = neglog_p), alpha = 0.9, shape = 21, color = 'black') +
  scale_fill_gradient(
    low = "#4C7CB4", high = "#CF6E67", name = expression(-log[10](adj.~p~value))
  ) +
  scale_size(range = c(3, 10), name = "Gene Ratio") +
  labs(
    x = "Combined Score",
    y = "Term"
  ) +
  theme_minimal(base_family = "Arial") +
  theme(
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = 'bold', family = "Arial"),
    panel.grid.major = element_line(color = "grey80"),  # faint grid lines
    panel.grid.minor = element_line(color = "grey90"),  # even fainter minor lines
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  # box around plot
    axis.line = element_line(color = "black", size = 0.8),     # <-- axis lines added
    axis.ticks = element_line(color = "black", size = 0.6)     # <-- optional: include ticks
  )

ggsave(gsoa1_plot,
       filename = "figures/goterms_overlap_pancreas_sig_genes_nkx2-3_bind_fMV_cap1_markers_method1.svg", 
       width = 10, 
       height = 5)


gsoa2 <- enrichAnnotateR(bMV_cap1_pancreas_sig_nkx_bind, database = 'function') %>%
    mutate(neglog_p = -log(Adjusted.P.value)) %>%
    filter(intersection_size > 2) %>%
  filter(term_size < 1000) %>%
  arrange(Combined.Score) %>%
  #slice_head(n = 20) %>% 
  mutate(Term = factor(Term, levels = Term))

write.csv(gsoa2, "output/Supplementary Table 7.csv", row.names = FALSE)


# Horizontal bar plot
gsoa2_plot <- ggplot(gsoa2, aes(x = Combined.Score,, y = Term)) +
  geom_point(aes(size = gene_ratio, color = neglog_p, fill = neglog_p), alpha = 0.9, shape = 21, color = 'black') +
  scale_fill_gradient(
    low = "#4C7CB4", high = "#CF6E67", name = expression(-log[10](adj.~p~value))
  ) +
  scale_size(range = c(3, 10), name = "Gene Ratio") +
  labs(
    x = "Combined Score",
    y = "Term"
  ) +
  theme_minimal(base_family = "Arial") +
  theme(
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = 'bold', family = "Arial"),
    panel.grid.major = element_line(color = "grey80"),  # faint grid lines
    panel.grid.minor = element_line(color = "grey90"),  # even fainter minor lines
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  # box around plot
    axis.line = element_line(color = "black", size = 0.8),     # <-- axis lines added
    axis.ticks = element_line(color = "black", size = 0.6)     # <-- optional: include ticks
  )

ggsave(gsoa2_plot,
       filename = "figures/goterms_overlap_pancreas_sig_genes_nkx2-3_bind_bMV_cap1_markers_method1.svg", 
       width = 10, 
       height = 5)



gsoa3 <- enrichAnnotateR(fMV_bMV_cap1, database = 'function') %>%
    mutate(neglog_p = -log(Adjusted.P.value)) %>%
    filter(intersection_size > 2) %>%
  filter(term_size < 1000) %>%
  arrange(Combined.Score) %>%
  #slice_head(n = 20) %>% 
  mutate(Term = factor(Term, levels = Term))


# Horizontal bar plot
gsoa3_plot <- ggplot(gsoa3, aes(x = Combined.Score,, y = Term)) +
  geom_point(aes(size = gene_ratio, color = neglog_p, fill = neglog_p), alpha = 0.9, shape = 21, color = 'black') +
  scale_fill_gradient(
    low = "#4C7CB4", high = "#CF6E67", name = expression(-log[10](adj.~p~value))
  ) +
  scale_size(range = c(3, 10), name = "Gene Ratio") +
  labs(
    x = "Combined Score",
    y = "Term"
  ) +
  theme_minimal(base_family = "Arial") +
  theme(
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = 'bold', family = "Arial"),
    panel.grid.major = element_line(color = "grey80"),  # faint grid lines
    panel.grid.minor = element_line(color = "grey90"),  # even fainter minor lines
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  # box around plot
    axis.line = element_line(color = "black", size = 0.8),     # <-- axis lines added
    axis.ticks = element_line(color = "black", size = 0.6)     # <-- optional: include ticks
  )

ggsave(gsoa3_plot,
       filename = "figures/goterms_overlap_fMV_cap1_bMV_cap1_markers_method1.svg", 
       width = 20, 
       height = 20)


gsoa1_terms <- as.vector(gsoa1$Term)
gsoa2_terms <- as.vector(gsoa2$Term)

intersect_genes <- intersect(gsoa1_terms, gsoa2_terms)
diff1 <- setdiff(gsoa1_terms, gsoa2_terms)
diff2 <- setdiff(gsoa2_terms, gsoa1_terms)


gsoa1_intersect <- gsoa1 %>%
  filter(Term %in% intersect_genes) %>%
  arrange(rev(Combined.Score))

# Horizontal bar plot
gsoa5_plot <- ggplot(gsoa1_intersect, aes(x = Combined.Score,, y = Term)) +
  geom_point(aes(size = gene_ratio, color = neglog_p, fill = neglog_p), alpha = 0.9, shape = 21, color = 'black') +
  scale_fill_gradient(
    low = "#4C7CB4", high = "#CF6E67", name = expression(-log[10](adj.~p~value))
  ) +
  scale_size(range = c(3, 10), name = "Gene Ratio") +
  labs(
    x = "Combined Score",
    y = "Term"
  ) +
  theme_minimal(base_family = "Arial") +
  theme(
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = 'bold', family = "Arial"),
    panel.grid.major = element_line(color = "grey80"),  # faint grid lines
    panel.grid.minor = element_line(color = "grey90"),  # even fainter minor lines
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  # box around plot
    axis.line = element_line(color = "black", size = 0.8),     # <-- axis lines added
    axis.ticks = element_line(color = "black", size = 0.6)     # <-- optional: include ticks
  )

ggsave(gsoa5_plot,
       filename = "figures/goterms_overlap_fMV_cap1_pancreas_nkx_bind_commonterms.svg", 
       width = 10, 
       height = 10)

gsoa2_intersect <- gsoa2 %>%
  filter(Term %in% intersect_genes) %>%
  arrange(rev(Combined.Score))

# Horizontal bar plot
gsoa6_plot <- ggplot(gsoa2_intersect, aes(x = Combined.Score,, y = Term)) +
  geom_point(aes(size = gene_ratio, color = neglog_p, fill = neglog_p), alpha = 0.9, shape = 21, color = 'black') +
  scale_fill_gradient(
    low = "#4C7CB4", high = "#CF6E67", name = expression(-log[10](adj.~p~value))
  ) +
  scale_size(range = c(3, 10), name = "Gene Ratio") +
  labs(
    x = "Combined Score",
    y = "Term"
  ) +
  theme_minimal(base_family = "Arial") +
  theme(
    axis.text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.title = element_text(size = 12, face = "bold", family = "Arial"),
    legend.text = element_text(size = 12, face = 'bold', family = "Arial"),
    panel.grid.major = element_line(color = "grey80"),  # faint grid lines
    panel.grid.minor = element_line(color = "grey90"),  # even fainter minor lines
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  # box around plot
    axis.line = element_line(color = "black", size = 0.8),     # <-- axis lines added
    axis.ticks = element_line(color = "black", size = 0.6)     # <-- optional: include ticks
  )

ggsave(gsoa6_plot,
       filename = "figures/goterms_overlap_bMV_cap1_pancreas_nkx_bind_commonterms.svg", 
       width = 15, 
       height = 15)

gsoa1_diff <- gsoa1 %>%
  filter(Term %in% diff1) %>%
  arrange(rev(Combined.Score))

gsoa2_diff <- gsoa2 %>%
  filter(Term %in% diff2) %>%
  arrange(rev(Combined.Score))

unique_sig_nkx_bMV <- setdiff(bMV_cap1_pancreas_sig_nkx_bind, fMV_cap1_pancreas_sig_nkx_bind)
unique_sig_nkx_fMV <- setdiff(fMV_cap1_pancreas_sig_nkx_bind, bMV_cap1_pancreas_sig_nkx_bind)
intersect_sig_nkx <- intersect(fMV_cap1_pancreas_sig_nkx_bind, bMV_cap1_pancreas_sig_nkx_bind)


# List all vectors
vecs <- list(
  "All Pancreas EC enriched genes" = signature_genes_all,
  "Pancreas EC enriched genes with NKX2-3 binding site in promoter" = pancreas_sig_nkx_bind,
  "fMV Capillary 1 markers" = markers_capillary_fmv_filt_genes,
  "bMV Capillary 1 markers" = markers_capillary_bmv_filt_genes,
  "Overlap of fMV Capillary 1 markers with Pancreas EC genes with NKX2-3 binding" = fMV_cap1_pancreas_sig_nkx_bind,
  "Overlap of bMV Capillary 1 markers with Pancreas EC genes with NKX2-3 binding" = bMV_cap1_pancreas_sig_nkx_bind,
  "Unique fMV Capillary 1 Markers that overlap with Pancreas EC genes with NKX2-3 binding" = unique_sig_nkx_fMV,
    "Unique bMV Capillary 1 Markers that overlap with Pancreas EC genes with NKX2-3 binding" = unique_sig_nkx_bMV,
  "Common fMV and bMV Capillary 1 Markers that overlap with Pancreas EC genes with NKX2-3 binding" = 
    intersect_sig_nkx
)

# Find max length
max_len <- max(sapply(vecs, length))

# Pad with NA
vecs_padded <- lapply(vecs, function(x) {
  length(x) <- max_len
  return(x)
})

# Create the data frame
df <- as.data.frame(vecs_padded, stringsAsFactors = FALSE)

# Save as CSV
write.csv(df, "output/Supplementary Table 5.csv", row.names = FALSE)

```

# Creating merged fMV_IlC pos neg object 
```{r}
# Load fMV data
fMV <- readRDS('output/fm_merge_analyzed_subset_analyzed/seurat_obj.rds')

# Update annotations if necessary
fMV@meta.data <- fMV@meta.data %>%
  mutate(cell_type = case_when(
    vascular_subtype == 'Arterial'  ~ "Capillary 3",
    vascular_subtype == 'Artery'  ~ "Arteriole",
    vascular_subtype == 'Venous'  ~ "Venule",
    vascular_subtype == 'Capillary 1'  ~ "Capillary 1",
    vascular_subtype == 'Capillary 2'  ~ "Capillary 2",
  ))

fMV@meta.data <- fMV@meta.data %>%
  mutate(dataset = case_when(
    dataset == 'fm_ilc_pos'  ~ "fMV+ILC",
    dataset == 'fm_only_pos'  ~ "fMV-Only",

  ))

fMV <- SetIdent(fMV, value = 'dataset')

fMV_ILC <- subset(fMV, value = 'fMV+ILC')

# Load ILC data
ilc <- readRDS('output/fm_ilc_neg_postqc_analyzed_subset_analyzed/seurat_obj.rds')

# Update annotations if necessary
ilc@meta.data <- ilc@meta.data %>%
  mutate(cell_type = case_when(
    leiden_res0_25 == '2'  ~ "Beta",
    leiden_res0_25 == '4'  ~ "Beta",
    leiden_res0_25 == '3'  ~ "Delta",
    leiden_res0_25 == '0'  ~ "Alpha",
    leiden_res0_25 == '1'  ~ "Alpha",
    leiden_res0_25 == '5'  ~ "Alpha",
  ))

# Subset only ILC clusters
ilc <- SetIdent(ilc, value = 'cell_type')

# Get metadata from ILC
ilc_meta <- ilc@meta.data

# Add ILC prefix to metadata rownames(Cells)
rownames(ilc_meta) <- paste0("ilc_", rownames(ilc_meta))
ilc_meta$barcode <- rownames(ilc_meta)
ilc_cells <- rownames(ilc_meta) 

# Get fMV metadata
fm_ilc_meta <- fMV_ILC@meta.data
# Add fm_ilc prefix to metadata frownames (Cells)
rownames(fm_ilc_meta) <- paste0("fm_", rownames(fm_ilc_meta))
fm_ilc_meta$barcode <- rownames(fm_ilc_meta)
fm_ilc_cells <- rownames(fm_ilc_meta) 

# Merge metas
merged_meta <- merge(fm_ilc_meta, ilc_meta, all = T)

# Order based on original order fMV first then ILC
merged_meta <- merged_meta[order(match(merged_meta$barcode, c(fm_ilc_cells, ilc_cells))), ]

# Convert human to mouse
# genes_m <- convertGenes(genes, convert_to = 'mouse', return_full = T)

# Get MAGIC imputed and normalized data and genes
magic_ilc <- Matrix(ilc@assays$MAGIC_imputed_data@layers$data, sparse = T)
genes_ilc <- rownames(ilc)
magic_fm_ilc <- Matrix(fMV_ILC@assays$MAGIC_imputed_data@layers$data, sparse = T)
genes_fm_ilc <- rownames(fMV_ILC)
# Convert fMV genes to human
genes_fm_ilc_h <- convertGenes(genes_fm_ilc)

# Get overlap of common genes
common_genes <- intersect(genes_ilc, genes_fm_ilc_h)

# Get indice of fMV gene ind
fm_ilc_gene_ind <- match(common_genes,genes_fm_ilc_h)

# Get indice of ILC gene ind
ilc_gene_ind <- match(common_genes, genes_ilc)

# Subset out only the common gene data
fm_ilc_magic_subset <- magic_fm_ilc[fm_ilc_gene_ind,]

# Subset out only the common gene data
ilc_magic_subset <- magic_ilc[ilc_gene_ind,]

# Merge only subset data
merge_data <- cbind(fm_ilc_magic_subset, ilc_magic_subset)

# 
rownames(merge_data) <- common_genes
colnames(merge_data) <- merged_meta$barcode

object <- CreateSeuratObject(counts = merge_data, data = merge_data, assay = 'RNA', meta.data = merged_meta)

saveRDS(object, file = 'output/fMV_ilc_neg_merged.rds')
```

# CellChat Pipeline
```{r}
data.input <- merge_data

object <- SetIdent(object, value = 'cell_type')

labels <- Idents(object)

meta <- data.frame(labels = labels, row.names = names(labels))

cellchat <- createCellChat(object = data.input, meta = meta, group.by = "labels")

CellChatDB <- CellChatDB.human
showDatabaseCategory(CellChatDB)
dplyr::glimpse(CellChatDB$interaction)

CellChatDB.use <- CellChatDB

cellchat@DB <- CellChatDB.use
cellchat <- subsetData(cellchat)

future::plan("multisession", workers = 4)
cellchat <- identifyOverExpressedGenes(cellchat)

cellchat <- identifyOverExpressedInteractions(cellchat)

cellchat <- computeCommunProb(cellchat, type = "triMean", trim = NULL, raw.use = TRUE)
```

# CellChat Pipeline Cont'd 
```{r}
cellchat <- filterCommunication(cellchat, min.cells = 10)

cellchat <- computeCommunProbPathway(cellchat)

cellchat <- aggregateNet(cellchat)

saveRDS(cellchat, file = "output/cellchat_merged_ilc_fm.rds")
```

CellChat: Load data
```{r}
cellchat <- readRDS(file = 'output/cellchat_merged_ilc_fm.rds')

# Color Pallete
color_pallete <- c("Arteriole" = '#B09C85', 
                   "Capillary 3" = '#EFC000',
                   "Venule" = '#FDAF91',
                   "Capillary 1" = '#0073C2',
                   "Capillary 2" = "#EE0000",
                   "Alpha" = "#4682B4",
                   "Beta" = '#B44656',
                   "Delta" = '#46B462'
                   )


```

# Circle Plot All Interactions
```{r}
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)

	png(filename = paste0("figures/all_interactions", ".png"), bg = "transparent", width = 00, height = 400)
	
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")

	dev.off()
```


# Circle Plot EC to ILC and ILC to EC
```{r}
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)


	png(filename = paste0("figures/ec_to_ilc", ".png"), bg = "transparent", width = 400, height = 400)
	par(mfrow = c(1, 1), xpd = TRUE)
	
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength", color.use = color_pallete, sources.use = 1:5, targets.use = 6:8)

	dev.off()
	
	png(filename = paste0("figures/ilc_to_ec", ".png"), bg = "transparent", width = 400, height = 400)
	par(mfrow = c(1, 1), xpd = TRUE)
	
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength", color.use = color_pallete, targets.use = 1:5, sources.use = 6:8)

	dev.off()
```

# Circle Plots of pathway of interest
```{r}
pdf("figures/ILC_to_EC_pathways.pdf", width = 14, height = 6, bg = "transparent")
par(mfrow = c(1, 6), mar = c(1, 1, 2, 1), oma = c(0, 0, 2, 0), xpd = TRUE)

# 2. First batch: to ECs (5 plots)
for (pathway in c("INSULIN", "CALCR", "SEMA3", "Testosterone", "PERIOSTIN")) {
  netVisual_aggregate(cellchat, signaling = pathway, layout = "circle", 
                      sources.use = 6:8, targets.use = 1:5, 
                      color.use = color_pallete)
  title(main = pathway, cex.main = 1.1)
}

dev.off()



pdf("figures/ECs_to_ILC_pathways.pdf", width = 14, height = 6, bg = "transparent")
par(mfrow = c(1, 6), mar = c(1, 1, 2, 1), oma = c(0, 0, 2, 0), xpd = TRUE)

# 3. Second batch: from ECs (6 plots)
for (pathway in c("TGFb", "SEMA7", "FGF", "HSPG", "ACTIVIN", "IGFBP")) {
  netVisual_aggregate(cellchat, signaling = pathway, layout = "circle",
                      color.use = color_pallete, 
                      sources.use = 1:5, targets.use = 6:10)
  title(main = pathway, cex.main = 1.1)
}

dev.off()



```

```{r}
pdf("figures/all_pathways_to_and_from_ECs.pdf", width = 14, height = 6, bg = "transparent")
par(mfrow = c(2, 6), mar = c(1, 1, 2, 1), oma = c(0, 0, 2, 0), xpd = TRUE)

# 1. Circle plot
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, 
                 weight.scale = TRUE, label.edge = FALSE, 
                 title.name = "Number of Interactions")

# 2. First batch: to ECs (5 plots)
for (pathway in c("INSULIN", "CALCR", "SEMA3", "Testosterone", "PERIOSTIN")) {
  netVisual_aggregate(cellchat, signaling = pathway, layout = "circle", 
                      sources.use = 6:8, targets.use = 1:5, 
                      color.use = color_pallete)
  title(main = pathway, cex.main = 1.1)
}

# 3. Second batch: from ECs (6 plots)
for (pathway in c("TGFb", "SEMA7", "FGF", "HSPG", "ACTIVIN", "IGFBP")) {
  netVisual_aggregate(cellchat, signaling = pathway, layout = "circle",
                      color.use = color_pallete, 
                      sources.use = 1:5, targets.use = 6:10)
  title(main = pathway, cex.main = 1.1)
}

dev.off()

```


# HeatMap
```{r}
# Access all the signaling pathways showing significant communications
pathways.show.all <- cellchat@netP$pathways

netVisual_heatmap(cellchat, signaling = pathways.show.all[10], color.heatmap = "Reds", color.use = color_pallete, targets.use = 1:5, sources.use = 6:10, remove.isolate = T, )
```

# Individual VEGF interactions
```{r}
pairLR <- extractEnrichedLR(cellchat, signaling = 'VEGF', geneLR.return = FALSE)
pairLR_plot <- pairLR[1:2,]

# Circle plot

for (i in 1:length(pairLR_plot)) {
	png(filename = paste0("figures/",pairLR_plot[i], ".png"), bg = "transparent", width = 400, height = 400)
		par(mfrow = c(1, 1), xpd = TRUE)
		
netVisual_individual(cellchat, signaling = 'VEGF', pairLR.use = pairLR_plot[i], layout = "circle", sources.use = 6:8, targets.use = 1:5, color.use = color_pallete)

	dev.off()
}

```

```{r}
LR.show <- as.data.frame(pairLR[1:2,]) # show one ligand-receptor pair

colnames(LR.show) <- 'interaction_name'

netVisual_bubble(cellchat, sources.use = 7, targets.use = c(1:5), remove.isolate = FALSE, pairLR.use = LR.show)
```


```{r}
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP") # the slot 'netP' means the inferred intercellular communication network of signaling pathways
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", signaling = pathways.show.all[1:20])
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", signaling = pathways.show.all[1:20])
ht1 + ht2
```

```{r}
netAnalysis_signalingRole_network(cellchat, signaling = 'VEGF', width = 8, height = 2.5, font.size = 10)
```

# RidgePlot individual interactions
```{r}

features1 <- c("VEGFA", "VEGFB", "VEGFC", "VEGFD")

features2 <- c("Flt1", "Kdr", "Flt4")

for (i in 1:length(features1)) {
  
  rPlot <- RidgePlot(neg_fMV, features = features1[i], group.by = 'cell_type', cols = color_pallete, assay = 'MAGIC_imputed_data')
  
  ggsave(rPlot, filename = paste0("RidgePlot_ILC_", features1[i], ".svg"), width = 4, height = 3)
  
}

for (i in 1:length(features2)) {
  
  rPlot <- RidgePlot(fMV_ILC, features = features2[i], group.by = 'cell_type', cols = color_pallete, assay = 'MAGIC_imputed_data')
  
  ggsave(rPlot, filename = paste0("RidgePlot_fMV_", features2[i], ".svg"), width = 5, height = 3)
  
}



features1 <- c("INS", "IAPP", "SEMA3A", "SRD5A1", "SEMA6D")

features2 <- c("Insr", "Calcrl", "Nrp1", "Ar", "Kdr")

for (i in 1:length(features1)) {
  
  rPlot <- RidgePlot(neg_fMV, features = features1[i], group.by = 'cell_type', cols = color_pallete, assay = 'MAGIC_imputed_data')
  
  ggsave(rPlot, filename = paste0("RidgePlot_ILC_", features1[i], ".svg"), width = 4, height = 3)
  
}

for (i in 1:length(features2)) {
  
  rPlot <- RidgePlot(fMV_ILC, features = features2[i], group.by = 'cell_type', cols = color_pallete, assay = 'MAGIC_imputed_data')
  
  ggsave(rPlot, filename = paste0("RidgePlot_fMV_", features2[i], ".svg"), width = 5, height = 3)
  
}



features1 <- c("TGFBR1", "ACVR1C","PLXNC1", "FGFR1")
features2 <- c("Tgfb1", "Inhbb", "Sema7a", "Fgf1")


for (i in 1:length(features1)) {
  
  rPlot <- RidgePlot(neg_fMV, features = features1[i], group.by = 'cell_type', cols = color_pallete, assay = 'MAGIC_imputed_data')
  
  ggsave(rPlot, filename = paste0("RidgePlot_ILC_", features1[i], ".svg"), width = 4, height = 3)
  
}

for (i in 1:length(features2)) {
  
  rPlot <- RidgePlot(fMV_ILC, features = features2[i], group.by = 'cell_type', cols = color_pallete, assay = 'MAGIC_imputed_data')
  
  ggsave(rPlot, filename = paste0("RidgePlot_fMV_", features2[i], ".svg"), width = 5, height = 3)
  
}
```
